---
title: "7-cell-calssification: auto+nabut"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Libraries
```{r}
library(MultiAssayExperiment)
library(SummarizedExperiment)
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(pheatmap)
library(ComplexHeatmap)
library(ggpubr)
library(reshape2)
library(PCAtools)
library(umap)
library(class)
library(cluster)
library(viridis)
library(boot)
library(EnhancedVolcano)
library(RColorBrewer)
```

## Functions
```{r}
source("~/single-cell-histone-ptm/functions.R")
```

## Load normalized data
```{r}
setwd("~/single-cell-histone-ptm/4-normalization/auto+nabut")

obj.filter.ratio.multi <- readRDS("obj.filter.ratio.multi.RDS")

# add unique replicate ids
obj.filter.ratio.multi$replicate.unique <- factor(as.numeric(factor(obj.filter.ratio.multi$experiment_replicate)))

# make ptm and proteins readable
rowData(obj.filter.ratio.multi[["ptm.ratio"]])$PTM.Name.Readable <- ptm_readable_translation(rownames(rowData(obj.filter.ratio.multi[["ptm.ratio"]])))

obj.filter.ratio.multi.1cell <- obj.filter.ratio.multi[,obj.filter.ratio.multi$multiplier == "1"]
```

## Visualize separation of conditions
- Since the treatment is nabut, this should cause higher global levels of acetylation

## Setup heatmap
```{r}
column_ha = HeatmapAnnotation(Replicate = obj.filter.ratio.multi.1cell$replicate.unique,
                              Condition = obj.filter.ratio.multi.1cell$condition,
                              col = list(Condition = c("control" = "#317ec2",
                                                       "treat" = "#c03830",
                                                       "mix" = "gray"),
                                           Replicate = c("1" = brewer.pal(5, "Set1")[1],
                                                        "2" = brewer.pal(5, "Set1")[2],
                                                        "3" = brewer.pal(5, "Set1")[3],
                                                        "4" = brewer.pal(5, "Set1")[4],
                                                        "5" = brewer.pal(5, "Set1")[5])))
```

### All PTMs
```{r}
dir.create("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")

# plot ratios
pdf("MS1.ratio.log1p.batchCorrect_all-PTM_heatmap.pdf",
    width = 20,
    height = 20)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = FALSE, scale = FALSE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Normalized Peptide Ratio"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = TRUE,
        split = rowData(obj.filter.ratio.multi.1cell[["ptm.ratio"]])$Protein.Name.Readable,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# plot Z-score
pdf("MS1.ratio.log1p.batchCorrect_all-PTM_scaled-heatmap.pdf",
    width = 10,
    height = 10)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = TRUE, scale = TRUE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        split = rowData(obj.filter.ratio.multi.1cell[["ptm.ratio"]])$Protein.Name.Readable,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# PCA
p <- pca(assay(obj.filter.ratio.multi.1cell[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect"),
         metadata = colData(obj.filter.ratio.multi.1cell),
         center = TRUE, 
         scale = TRUE)

pdf("MS1.ratio.log1p.batchCorrect_all-PTM_screeplot.pdf",
    width = 5,
    height = 5)
screeplot(p,
          components = getComponents(p)[1:10])
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_all-PTM_eigencor.pdf",
    width = 5,
    height = 5)
eigencorplot(p,
             main = "Principle component correlations",
             cexMain = 1.5,
             metavars = c("replicate.unique", "date", "run_id", "experiment", "condition", "total.ms1"),
             col = viridis(100),
             colCorval = 'firebrick',
             fontCorval = 2,
             cexCorval = 0.5,
             rotLabX = 45,
             posColKey = 'top')
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_all-PTM_pca.pdf",
    width = 5,
    height = 5)
biplot(p,
       colby = "condition", 
       #shape = "replicate.unique",
       legendPosition = 'top',
       hline = 0, 
       vline = 0,
       lab = NULL) +
  labs(color = "Condition",
       shape = "Replicate") +
  theme_Publication() +
  theme(legend.box="vertical", 
        legend.margin=margin()) +
  scale_color_manual(values = c("control" = "#317ec2",
                                "treat" = "#c03830",
                                "mix" = "gray"),
                     labels = c("Control",
                                "Mixed",
                                "Treated"))
dev.off()
```

### Top 25% variable PTMs
```{r}
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")

# get top 25% variable
top25 <- names(sort(rowVars(assay(obj.filter.ratio.multi.1cell[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect")), decreasing = TRUE))[1:(ceiling(nrow(obj.filter.ratio.multi.1cell[["ptm.ratio"]])*0.25))]

df <- assay(obj.filter.ratio.multi.1cell[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect")[rownames(obj.filter.ratio.multi.1cell[["ptm.ratio"]]) %in% top25,]

histone_proteins <- sub("^([^-]+)-.*$", "\\1", rownames(df))

# plot ratios
pdf("MS1.ratio.log1p.batchCorrect_top25_heatmap.pdf",
    width = 10,
    height = 5)
Heatmap(t(scale(t(df), center = FALSE, scale = FALSE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Normalized Peptide Ratio"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# plot Z-score
pdf("MS1.ratio.log1p.batchCorrect_top25_scaled-heatmap.pdf",
    width = 10,
    height = 5)
Heatmap(t(scale(t(df), center = TRUE, scale = TRUE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()
```

### Global PTMs
```{r}
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")

# plot ratios
pdf("MS1.global.ratio.batchCorrect_heatmap.pdf",
    width = 8,
    height = 3)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["global.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = FALSE, scale = FALSE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Normalized Peptide Ratio"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        #split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# plot Z-score
pdf("MS1.global.ratio.batchCorrect_scaled-heatmap.pdf",
    width = 8,
    height = 3)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["global.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = TRUE, scale = TRUE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        #split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()
```

### H4 ac
```{r}
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")

# plot ratios
rownames(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]]) <- c("0-ac", "Mono-ac", "Di-ac", "Tri-ac", "Tetra-ac")

pdf("MS1.h4.ac.ratio.batchCorrect_heatmap.pdf",
    width = 8,
    height = 3)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = FALSE, scale = FALSE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Normalized Peptide Ratio"),
        cluster_rows = FALSE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        #split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# plot Z-score
pdf("MS1.h4.ac.ratio.batchCorrect_scaled-heatmap.pdf",
    width = 8,
    height = 3)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = TRUE, scale = TRUE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = FALSE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        #split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# PCA
p <- pca(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect"),
         metadata = colData(obj.filter.ratio.multi.1cell),
         center = TRUE, 
         scale = TRUE,
         removeVar = 0)

pdf("MS1.h4.ac.ratio.log1p.batchCorrect_all-PTM_eigencor.pdf",
    width = 5,
    height = 5)
eigencorplot(p,
             getComponents(p, seq_len(5)),
             main = "Principle component correlations",
             cexMain = 1.5,
             metavars = c("replicate.unique", "date", "run_id", "experiment", "condition", "total.ms1"),
             col = viridis(100),
             colCorval = 'firebrick',
             fontCorval = 2,
             cexCorval = 0.5,
             rotLabX = 45,
             posColKey = 'top')
dev.off()

pdf("MS1.h4.ac.ratio.batchCorrect_pca.pdf",
    width = 5,
    height = 5)
biplot(p,
       colby = "condition", 
       #shape = "replicate.unique",
       legendPosition = 'top',
       hline = 0, 
       vline = 0,
       lab = NULL) +
  labs(color = "Condition",
       shape = "Replicate") +
  theme_Publication() +
  theme(legend.box="vertical", 
        legend.margin=margin()) +
  scale_color_manual(values = c("control" = "#317ec2",
                                "treat" = "#c03830",
                                "mix" = "gray"),
                     labels = c("Control",
                                "Mixed",
                                "Treated"))
dev.off()
```


## KNN class training and prediction
- Train a knn classifer on the PTMs found to be significantly different between control and nabut
- Only use single cells as bulk samples may distort classification

#### Training
```{r}
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")

# split training into 50% of data (control and treat)
control_names <- rownames(colData(obj.filter.ratio.multi.1cell)[obj.filter.ratio.multi.1cell$condition == "control",])
treat_names <- rownames(colData(obj.filter.ratio.multi.1cell)[obj.filter.ratio.multi.1cell$condition == "treat",])

# Calculate the size of 75% of the vector
size_50_percent_control <- round(length(control_names) * 0.5)
size_50_percent_treat <- round(length(treat_names) * 0.5)

# Randomly select x% of the names for training
set.seed(123)
control_names_training <- sample(control_names, size = size_50_percent_control)
treat_names_training <- sample(treat_names, size = size_50_percent_treat)

# Assign the remaining x% of the names for testing
control_names_testing <- setdiff(control_names, control_names_training)
treat_names_testing <- setdiff(treat_names, treat_names_training)

# Assign labels of each condition
train_condition <- factor(colData(obj.filter.ratio.multi.1cell)[colnames(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]]) %in% c(control_names_training, treat_names_training),]$condition)
test_condtion <- factor(colData(obj.filter.ratio.multi.1cell)[colnames(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]]) %in% c(control_names_testing, treat_names_testing),]$condition)

# get scaled h4 data
train <- scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")[,c(control_names_training, treat_names_training)]))
test <- scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")[,c(control_names_testing, treat_names_testing)]))

# eval prediction model
knn.pred.known <- knn(train = train, 
                test = test,
                cl = train_condition, 
                k = 2)
names(knn.pred.known) <- c(control_names_testing, treat_names_testing)

res <- table(test_condtion, knn.pred.known)
accuracy <- sum(diag(res))/length(test_condtion)
print(accuracy)

# assign results
df <- colData(obj.filter.ratio.multi.1cell) %>%
  as.data.frame() %>%
  rownames_to_column("row_names") %>% # Convert rownames to a column for matching
  mutate(
    knn.training.split = case_when(row_names %in% c(control_names_training, treat_names_training) ~ "train", 
                          row_names %in% c(control_names_testing, treat_names_testing) ~ "test",
                          TRUE ~ NA_character_))
obj.filter.ratio.multi.1cell$knn.training.split <- df$knn.training.split

# plot
column_ha = HeatmapAnnotation("KNN Training" = as.factor(obj.filter.ratio.multi.1cell$knn.training.split),
                              Condition = obj.filter.ratio.multi.1cell$condition,
                               col = list("KNN Training" = c("train" = brewer.pal(3, "Set2")[1],
                                                             "test" = brewer.pal(3, "Set2")[2]),
                                          Condition = c("control" = "#317ec2",
                                                       "treat" = "#c03830",
                                                       "mix" = "gray")
                                          )
                              )

pdf("MS1.h4.ac.ratio.batchCorrect_knn-training_scaled-heatmap.pdf",
    width = 8,
    height = 3)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = TRUE, scale = TRUE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = FALSE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        #split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()
```


#### Prediction
```{r}
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")

# Split data into known and unknown
train_names <- colnames(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]])[obj.filter.ratio.multi.1cell$condition != "mix"]
test_names <- colnames(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]])[obj.filter.ratio.multi.1cell$condition == "mix"]

# Assign labels of each condition
train_labels <- factor(colData(obj.filter.ratio.multi.1cell)[colnames(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]]) %in% train_names,]$condition)

# get scaled h4 data
train <- scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")[,train_names]))
test <- scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")[,test_names]))

# eval prediction model
knn.pred.mix <- knn(train = train, 
                test = test,
                cl = train_labels, 
                k = 2)
# results
names(knn.pred.mix) <- test_names
sum(knn.pred.mix == "control")
sum(knn.pred.mix == "treat")

# assign results
df <- colData(obj.filter.ratio.multi.1cell) %>%
  as.data.frame() %>%
  rownames_to_column("row_names") %>% # Convert rownames to a column for matching
  mutate(
    knn.mix.predict = case_when(row_names %in%  names(knn.pred.mix)[knn.pred.mix == "control"] ~ "control", 
                          row_names %in% names(knn.pred.mix)[knn.pred.mix == "treat"] ~ "treat",
                          TRUE ~ NA_character_))
obj.filter.ratio.multi.1cell$knn.mix.predict <- df$knn.mix.predict

# make new condition column that merges known and predicted labels
condition.temp <- as.character(obj.filter.ratio.multi.1cell$condition)
condition.temp[condition.temp == "mix"] <- NA
obj.filter.ratio.multi.1cell$condition.predict <- coalesce(condition.temp, obj.filter.ratio.multi.1cell$knn.mix.predict)

# heatmap
column_ha = HeatmapAnnotation(Batch = obj.filter.ratio.multi.1cell$replicate.unique,
                              "Known Condition" = obj.filter.ratio.multi.1cell$condition,
                              "KNN Training" = as.factor(obj.filter.ratio.multi.1cell$knn.training.split),
                              "KNN Prediction" = as.factor(obj.filter.ratio.multi.1cell$knn.mix.predict),
                               col = list(Batch = c("1" = brewer.pal(5, "Set1")[1],
                                                        "2" = brewer.pal(5, "Set1")[2],
                                                        "3" = brewer.pal(5, "Set1")[3],
                                                        "4" = brewer.pal(5, "Set1")[4],
                                                        "5" = brewer.pal(5, "Set1")[5]),
                                          "KNN Prediction" = c("control" = "#317ec2",
                                                               "treat" = "#c03830"),
                                          "KNN Training" = c("train" = brewer.pal(3, "Set2")[1],
                                                             "test" = brewer.pal(3, "Set2")[2]),
                                          "Known Condition" = c("control" = "#317ec2",
                                                       "treat" = "#c03830",
                                                       "mix" = "gray")
                                          ),
                              na_col = "white"
                              )

pdf("MS1.h4.ac.ratio.batchCorrect_knn-predict_scaled-heatmap.pdf",
    width = 8,
    height = 3)
Heatmap(t(scale(t(assay(obj.filter.ratio.multi.1cell[["h4.ac.ratio"]], "MS1.ratio.log1p.batchCorrect")), center = TRUE, scale = TRUE)),
        top_annotation = column_ha,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = FALSE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        #split = histone_proteins,
        cluster_column_slices = TRUE,
        cluster_row_slices = TRUE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# PCA
p <- pca(assay(obj.filter.ratio.multi.1cell[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect"),
         metadata = colData(obj.filter.ratio.multi.1cell),
         center = TRUE, 
         scale = TRUE)

pdf("MS1.h4.ac.ratio.batchCorrect_knn-predict_pca.pdf",
    width = 4,
    height = 5)
biplot(p,
       colby = "condition.predict", 
       legendPosition = 'top',
       hline = 0, 
       vline = 0,
       lab = NULL,
       encircle = TRUE,
       encircleFillKey = c("control" = "#317ec2",
                                "treat" = "#c03830"),
       encircleAlpha = 0.2) +
  labs(color = "Condition") +
  theme_Publication() +
  theme(legend.box="vertical", 
        legend.margin=margin()) +
  scale_color_manual(values = c("control" = "#317ec2",
                                "treat" = "#c03830"),
                     labels = c("Control",
                                "Treated"))
dev.off()
```

#### Save object
```{r}
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut/without_100-cells")

saveRDS(obj.filter.ratio.multi.1cell, "obj.filter.ratio.multi.1cell.RDS")
```

# Integrate KNN without into full object
```{r}
df <- colData(obj.filter.ratio.multi.1cell) %>%
  as.data.frame() %>%
  rownames_to_column("row_names") %>% # Convert rownames to a column for matching
  mutate(
    knn.mix.predict = case_when(row_names %in%  names(knn.pred.mix)[knn.pred.mix == "control"] ~ "control", 
                          row_names %in% names(knn.pred.mix)[knn.pred.mix == "treat"] ~ "treat",
                          TRUE ~ NA_character_))
obj.filter.ratio.multi.1cell$knn.mix.predict <- df$knn.mix.predict

# make new condition column that merges known and predicted labels
condition.temp <- as.character(obj.filter.ratio.multi.1cell$condition)
condition.temp[condition.temp == "mix"] <- NA
obj.filter.ratio.multi.1cell$condition.predict <- coalesce(condition.temp, obj.filter.ratio.multi.1cell$knn.mix.predict)

# PCA
p <- pca(assay(obj.filter.ratio.multi.1cell[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect"),
         metadata = colData(obj.filter.ratio.multi.1cell),
         center = TRUE, 
         scale = TRUE)

pdf("MS1.ptm.ratio.batchCorrect_knn-predict_pca_integrated.pdf",
    width = 4,
    height = 5)
biplot(p,
       colby = "condition.predict", 
       legendPosition = 'top',
       hline = 0, 
       vline = 0,
       lab = NULL,
       encircle = TRUE,
       encircleFillKey = c("control" = "#317ec2",
                           "treat" = "#c03830"),
       encircleAlpha = 0.2) +
  labs(color = "Condition") +
  theme_Publication() +
  theme(legend.box="vertical", 
        legend.margin=margin()) +
  scale_color_manual(values = c("control" = "#317ec2",
                                "treat" = "#c03830"),
                     labels = c("Control",
                                "Treated"))
dev.off()
```

## Save integrated object
```{r}
setwd("~/single-cell-histone-ptm/8-cell-classification/auto+nabut")

saveRDS(obj.filter.ratio.multi, "obj.filter.ratio.multi.RDS")
```
