---
title: "2-derivitization-efficiency"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Libraries
```{r}
library(SummarizedExperiment)
library(MultiAssayExperiment)
library(PCAtools)
library(ComplexHeatmap)
library(BatchQC)
library(sva)
library(tidyr)
library(dplyr)
library(tibble)
library(purrr)
library(ggplot2)
library(pheatmap)
library(ggpubr)
library(reshape2)
library(viridis)
library(ggrepel)
library(missForest)
library(limma)
library(stringr)
library(scales)
```

## Functions
```{r}
source("~/single-cell-histone-ptm/functions.R")
```

## Load data
```{r}
# metadata
setwd("~/single-cell-histone-ptm/1-data/single-cell")

coldata <- read.csv("sample-table.csv", 
               header = TRUE,
               stringsAsFactors = TRUE)

# skyline output
setwd("~/single-cell-histone-ptm/1-data/single-cell/intensities")

# combine all plates
df.combine <- read.csv("241017_Skyline_SC_export_auto1_auto2_auto4_nabut1_nabut4_prop_efficiency.csv", 
               header = TRUE,
               stringsAsFactors = TRUE,
               na.strings = "#N/A")

# combine into summarized experiment object
obj <- load_summarized_experiment_prop_eff_skyline(df.combine, coldata)

# add total MS1 and MS2 to coldata
obj$total.ms1 <- colSums(assay(obj, "MS1"), na.rm = TRUE)
obj$total.ms2 <- colSums(assay(obj, "MS2"), na.rm = TRUE)

# change replicate 4 to 3 to make it easier to read for auto
# change replicate 4 to 2 to make it easier to read for nabut
obj$replicate <- as.character(obj$replicate)
obj$replicate[obj$replicate == 4 & obj$experiment == "auto"] <- 3
obj$replicate[obj$replicate == 4 & obj$experiment == "nabut"] <- 2
obj$replicate <- factor(obj$replicate)

# set cells from auto experiments to "control" condition - these were not treated
obj$condition[obj$experiment == "auto"] <- "control"
```

# Calculate propionylation efficiency
- This is a ratio between the intensity of the fully propionylated peptide to the under or over propionylated peptides

## Under propionylation in 1st round - tryptic peptides with propionlyation modifications on lysine and peptide N-termini
```{r}
obj$prop.eff.underprop.first.round <- unlist(1 - assay(obj, "MS1")["Full prop (ArgC)",]/
  colSums(assay(obj, "MS1")[c("Full prop (ArgC)", "1st round underprop (fully tryptic)"),]))
```

## Under propionylation in 2nd round - argC peptides with no propionlyation modifications on peptide N-termini
```{r}
obj$prop.eff.underprop.second.round <- unlist(1 - assay(obj, "MS1")["Full prop (ArgC)",]/
  colSums(assay(obj, "MS1")[c("Full prop (ArgC)", "2nd round underprop (ArgC)"),]))
```

## Under propionylation in 1st and 2nd round - tryptic peptides with no propionylation modifications
```{r}
obj$prop.eff.underprop.both.round <- unlist(1 - assay(obj, "MS1")["Full prop (ArgC)",]/
  colSums(assay(obj, "MS1")[c("Full prop (ArgC)", "No prop (fully tryptic)"),]))
```

## Under propionylation total - sum of all 3 types of underpropionylation
```{r}
obj$prop.eff.underprop.summary <- unlist(1 - assay(obj, "MS1")["Full prop (ArgC)",]/
  colSums(assay(obj, "MS1")[c("Full prop (ArgC)", 
                              "No prop (fully tryptic)", 
                              "1st round underprop (fully tryptic)", 
                              "2nd round underprop (ArgC)"),]))
```

## Over propionylation - argC peptides with propionlyation modifications on lysine and peptide N-termini and on tyrosine
```{r}
obj$prop.eff.overprop <- unlist(1 - assay(obj, "MS1")["Full prop (ArgC)",]/
  colSums(assay(obj, "MS1")[c("Full prop (ArgC)","Overprop on T (ArgC)"),]))
```

## Output propionylation efficiency ratios
```{r}
setwd("~/single-cell-histone-ptm/2-derivitization-efficiency/DIA")

df <- as.data.frame(colData(obj))
write.csv(df[, grep("prop", colnames(df))], "propionylation_efficiency_percent.csv")
```

## Save object
```{r}
setwd("~/single-cell-histone-ptm/2-derivitization-efficiency/DIA")

saveRDS(obj, "obj.RDS")
```

# Plot propionylation efficiency

## Auto cells

### All runs
```{r}
setwd("~/single-cell-histone-ptm/2-derivitization-efficiency/DIA")

df <- colData(obj) %>% 
  as.data.frame() %>%
  rownames_to_column("run") %>%
  pivot_longer(
    cols = starts_with("prop"),
    names_to = "variable",
    values_to = "value") %>%
  filter(experiment == "auto" & multiplier != 0)
df$variable <- factor(df$variable,
                      levels = c("prop.eff.underprop.summary",
                                 "prop.eff.overprop","prop.eff.underprop.first.round",
                                 "prop.eff.underprop.second.round",
                                 "prop.eff.underprop.both.round"))

# summary
pdf("auto_cells_propionylation_efficiency_summary.pdf",
    width = 6,
    height = 5)
ggplot(subset(df, !grepl("round", variable)), aes(x = variable, y = value)) +
  geom_boxplot(aes(fill = replicate), outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(aes(fill = replicate, shape = as.factor(multiplier)), 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75, jitter.height = 0), 
              alpha = 0.75, size = 1.5) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_discrete(labels = c("Under\nPropionylation", "Over\nPropionlyation")) +
  labs(x = "", 
       shape = "# Cells Per Injection",
       y = "% Raw MS1 Area of Peptide\nNot Correctly Propionylated", 
       fill = "Batch") +
  theme_Publication() +
  theme(legend.box = "vertical")
dev.off()

# underpropionylation details
pdf("auto_cells_propionylation_efficiency_underpropionylated.pdf",
    width = 6,
    height = 5)
ggplot(subset(df, grepl("round", variable)), aes(x = variable, y = value)) +
  geom_boxplot(aes(fill = replicate), outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(aes(fill = replicate, shape = as.factor(multiplier)), 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75, jitter.height = 0), 
              alpha = 0.75, size = 1.5) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_discrete(labels = c("1st Round", "2nd Round", "1st & 2nd Rounds")) +
  labs(x = "", 
       shape = "# Cells Per Injection",
       y = "% Raw MS1 Area of Peptide\nUnder Propionylated", 
       fill = "Batch") +
  theme_Publication() +
  theme(legend.box = "vertical")
dev.off()
```

### Single cells
```{r}
setwd("~/single-cell-histone-ptm/2-derivitization-efficiency/DIA")

df <- colData(obj) %>% 
  as.data.frame() %>%
  rownames_to_column("run") %>%
  pivot_longer(
    cols = starts_with("prop"),
    names_to = "variable",
    values_to = "value") %>%
  filter(experiment == "auto" & multiplier == 1)
df$variable <- factor(df$variable,
                      levels = c("prop.eff.underprop.summary",
                                 "prop.eff.overprop","prop.eff.underprop.first.round",
                                 "prop.eff.underprop.second.round",
                                 "prop.eff.underprop.both.round"))

# summary
pdf("auto_cells_1cell_propionylation_efficiency_summary.pdf",
    width = 6,
    height = 5)
ggplot(subset(df, !grepl("round", variable)), aes(x = variable, y = value)) +
  geom_boxplot(aes(fill = replicate), outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(aes(fill = replicate), 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75, jitter.height = 0), 
              alpha = 0.75, size = 1.5) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_discrete(labels = c("Under\nPropionylation", "Over\nPropionlyation")) +
  labs(x = "", 
       shape = "# Cells Per Injection",
       y = "% Raw MS1 Area of Peptide\nNot Correctly Propionylated", 
       fill = "Batch") +
  theme_Publication() +
  theme(legend.box = "vertical")
dev.off()

# underpropionylation details
pdf("auto_cells_1cell_propionylation_efficiency_underpropionylated.pdf",
    width = 6,
    height = 5)
ggplot(subset(df, grepl("round", variable)), aes(x = variable, y = value)) +
  geom_boxplot(aes(fill = replicate), outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(aes(fill = replicate), 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75, jitter.height = 0), 
              alpha = 0.75, size = 1.5) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_discrete(labels = c("1st Round", "2nd Round", "1st & 2nd Rounds")) +
  labs(x = "", 
       shape = "# Cells Per Injection",
       y = "% Raw MS1 Area of Peptide\nUnder Propionylated", 
       fill = "Batch") +
  theme_Publication() +
  theme(legend.box = "vertical")
dev.off()
```

## Nabut cells
- These are single cells only
```{r}
setwd("~/single-cell-histone-ptm/2-derivitization-efficiency/DIA")

df <- colData(obj) %>% 
  as.data.frame() %>%
  rownames_to_column("run") %>%
  pivot_longer(
    cols = starts_with("prop"),
    names_to = "variable",
    values_to = "value") %>%
  filter(experiment == "nabut" & multiplier == 1)
df$variable <- factor(df$variable,
                      levels = c("prop.eff.underprop.summary",
                                 "prop.eff.overprop","prop.eff.underprop.first.round",
                                 "prop.eff.underprop.second.round",
                                 "prop.eff.underprop.both.round"))

# summary
pdf("nabut_cells_1cell_propionylation_efficiency_summary.pdf",
    width = 4,
    height = 5)
ggplot(subset(df, !grepl("round", variable)), aes(x = variable, y = value)) +
  geom_boxplot(aes(fill = replicate), outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(aes(fill = replicate), 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75, jitter.height = 0), 
              alpha = 0.75, size = 1.5) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_discrete(labels = c("Under\nPropionylation", "Over\nPropionlyation")) +
  labs(x = "", 
       shape = "# Cells Per Injection",
       y = "% Raw MS1 Area of Peptide\nNot Correctly Propionylated", 
       fill = "Batch") +
  theme_Publication() +
  theme(legend.box = "vertical")
dev.off()

# underpropionylation details
pdf("nabut_cells_1cell_propionylation_efficiency_underpropionylated.pdf",
    width = 4,
    height = 5)
ggplot(subset(df, grepl("round", variable)), aes(x = variable, y = value)) +
  geom_boxplot(aes(fill = replicate), outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(aes(fill = replicate), 
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75, jitter.height = 0), 
              alpha = 0.75, size = 1.5) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_x_discrete(labels = c("1st Round", "2nd Round", "1st & 2nd Rounds")) +
  labs(x = "", 
       shape = "# Cells Per Injection",
       y = "% Raw MS1 Area of Peptide\nUnder Propionylated", 
       fill = "Batch") +
  theme_Publication() +
  theme(legend.box = "vertical")
dev.off()
```