---
title: "4-sc-vs-bulk: auto"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Libraries
```{r}
library(SummarizedExperiment)
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(pheatmap)
library(ComplexHeatmap)
library(ggpubr)
library(reshape2)
library(PCAtools)
library(viridis)
library(stringr)
```

## Functions
```{r}
source("~/single-cell-histone-ptm/Analysis/functions.R")
```

## Load data

### Single cells and 100 cell bulk
```{r}
setwd("~/single-cell-histone-ptm/Analysis/4-normalization/auto")

obj <- readRDS("obj.filter.RDS")

obj.filter.ratio.multi <- readRDS("obj.filter.ratio.multi.RDS")
```

### Bulk from extracted histones samples 
- singlePTM data
- from using -50 and -70 CV. Needs to be extracted from -50 though
```{r}
setwd("~/single-cell-histone-ptm/Analysis/1-data/bulk/230816_nabut_bulk_test")

nabut_bulk_230816 <- read.csv("histone_ratios_single_PTMs.csv", header = TRUE, row.names = 1)
rownames(nabut_bulk_230816) <- sub("(H\\d+)(K\\d+)([a-z]+\\d*)", "\\1-\\2[\\3]", rownames(nabut_bulk_230816))

setwd("~/single-cell-histone-ptm/Analysis/1-data/bulk/231012_broad")

nabut_bulk_231012 <- read.csv("histone_ratios_single_PTMs.csv", header = TRUE, row.names = 1)
rownames(nabut_bulk_231012) <- sub("(H\\d+)(K\\d+)([a-z]+\\d*)", "\\1-\\2[\\3]", rownames(nabut_bulk_231012))
```

### Bulk from Simone's previous studes
```{r}
setwd("~/single-cell-histone-ptm/Analysis/1-data/bulk/simone")

bottomup <- read.csv("Bottom-up vs middle-down (Anal Chem 2015) - cleaned data.csv", header = TRUE)

dialowres <- read.csv("DIA low resolution (Anal Chem 2015) - cleaned data.csv", header = TRUE)
```

#### Calculate single PTMs
```{r}
bottomup.sptm <- bottomup %>%
  mutate(sptm = split_modifications(Peptide.Note)) %>% # split the ptm into single ptms
  unnest(sptm) %>% # make each sptm into it's own row
  separate(col = sptm, into = c("aa", "mod"), sep = "(?<=[0-9])(?=[a-zA-Z])", remove = TRUE, convert = FALSE) %>% # separate into aa residue and type of mod
  group_by(Protein, aa, mod) %>% # group 
  summarise(across(where(is.numeric), sum), .groups = 'drop') %>% # sum ratios for aa mods across peptides 
  mutate(sptm = paste0(Protein, "-", aa, "[", mod, "]")) %>%
  column_to_rownames(var = "sptm") %>%
  select(-c(Protein, aa, mod))

dialowres.sptm <- dialowres %>%
  mutate(sptm = split_modifications(Peptide.Note)) %>% # split the ptm into single ptms
  unnest(sptm) %>% # make each sptm into it's own row
  separate(col = sptm, into = c("aa", "mod"), sep = "(?<=[0-9])(?=[a-zA-Z])", remove = TRUE, convert = FALSE) %>% # separate into aa residue and type of mod
  group_by(Protein, aa, mod) %>% # group 
  summarise(across(where(is.numeric), sum), .groups = 'drop') %>% # sum ratios for aa mods across peptides 
  mutate(sptm = paste0(Protein, "-", aa, "[", mod, "]")) %>%
  column_to_rownames(var = "sptm") %>%
  select(-c(Protein, aa, mod))

# match to single cell sptm
# bottomup.sptm.filter <- bottomup.sptm[rownames(bottomup.sptm) %in% rownames(obj.filter.ratio.multi[["sptm.ratio"]]),]
# dialowres.sptm.filter <- dialowres.sptm[rownames(dialowres.sptm) %in% rownames(obj.filter.ratio.multi[["sptm.ratio"]]),]
```

# Bulk vs SC

## Log transform

### MS1
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

# correlation all replicates
df <- assay(obj.filter.ratio.multi[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj)), by.x = "variable", by.y = "row.names") %>%
  group_by(multiplier, ptm) %>%
  summarize(pseudobulk = log2(sum(value, na.rm = TRUE)),
            bulk = log2(mean(value, na.rm = TRUE)))

df.pb <- data.frame(pseudobulk = subset(df, multiplier == 1)$pseudobulk,
                    bulk = subset(df, multiplier == 100)$bulk,
                    row.names = subset(df, multiplier == 1)$ptm)

pdf("MS1.log2.sc_vs_bulk.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk)) +
  geom_point() +
  stat_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  labs(x = "Bulk\nLog2 Mean MS1 Area",
       y = "Pseudobulk\nLog2 Summed MS1 Area")
dev.off()

# correlation by replicate
df <- assay(obj.filter.ratio.multi[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
    rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj)), by.x = "variable", by.y = "row.names") %>%
  group_by(multiplier, ptm, replicate) %>%
  summarize(pseudobulk = log2(sum(2^value, na.rm = TRUE)),
            bulk = log2(mean(2^value, na.rm = TRUE)))

df.pb <- data.frame(pseudobulk = subset(df, multiplier == 1)$pseudobulk,
                    bulk = subset(df, multiplier == 100)$bulk,
                    replicate = subset(df, multiplier == 1)$replicate,
                    ptm = subset(df, multiplier == 1)$ptm)

pdf("MS1.log2.sc_vs_bulk_replicate_no_cor.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk, color = as.factor(replicate))) +
  geom_point() +  # Adds points
  stat_smooth(method = "lm") +  # Adds a linear model fit
  scale_color_discrete(name = "Replicate") +  # Names the color legend as "Replicate"
  labs(x = "Bulk\nMean Log2(MS1 Intensity)", y = "Pseudobulk\nSummed Log2(MS1 Intensity)") +
  theme_Publication(base_size = 16) +  # Custom theme
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  theme(legend.position = "top") +
  guides(color = guide_legend(override.aes = list(shape = 19)))
dev.off()

pdf("MS1.log2.sc_vs_bulk_replicate_cor.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk, color = as.factor(replicate))) +
  geom_point() +  # Adds points
  stat_smooth(method = "lm") +  # Adds a linear model fit
  stat_cor() +
  scale_color_discrete(name = "Replicate") +  # Names the color legend as "Replicate"
  labs(x = "Bulk\nMean Log2(MS1 Intensity)", y = "Pseudobulk\nSummed Log2(MS1 Intensity)") +
  theme_Publication(base_size = 16) +  # Custom theme
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  theme(legend.position = "top") +
  guides(color = guide_legend(override.aes = list(shape = 19)))
dev.off()

pdf("MS1.log2.sc_vs_bulk_replicate_cor_transparent.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk, color = as.factor(replicate))) +
  geom_point() +  # Adds points
  stat_smooth(method = "lm") +  # Adds a linear model fit
  stat_cor() +
  scale_color_discrete(name = "Replicate") +  # Names the color legend as "Replicate"
  labs(x = "Bulk\nMean Log2(MS1 Intensity)", y = "Pseudobulk\nSummed Log2(MS1 Intensity)") +
  theme_Publication(base_size = 16) +  # Custom theme
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
        theme(legend.position = "top",
        plot.background = element_rect(fill = "transparent", colour = NA), # transparent background
        panel.background = element_rect(fill = "transparent", colour = NA), # transparent panel
        legend.background = element_rect(fill = "transparent", colour = NA), # transparent legend
        legend.key = element_rect(fill = "transparent", colour = NA)) + # transparent legend keys)
  guides(color = guide_legend(override.aes = list(shape = 19)))
dev.off()
```

### MS2
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

# correlation all replicates
df <- assay(obj, "MS2.log2.batchCorrect") %>% 
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj)), by.x = "variable", by.y = "row.names") %>%
  group_by(multiplier, ptm) %>%
  summarize(pseudobulk = log2(sum(2^value, na.rm = TRUE)),
            bulk = log2(mean(2^value, na.rm = TRUE)))

df.pb <- data.frame(pseudobulk = subset(df, multiplier == 1)$pseudobulk,
                    bulk = subset(df, multiplier == 100)$bulk,
                    row.names = subset(df, multiplier == 1)$ptm)

pdf("MS2.log2.sc_vs_bulk.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk)) +
  geom_point() +
  stat_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  labs(x = "Bulk\nLog2 Mean MS2 Area",
       y = "Pseudobulk\nLog2 Summed MS2 Area")
dev.off()

# correlation by replicate
df <- assay(obj, "MS2.log2.batchCorrect") %>% 
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj)), by.x = "variable", by.y = "row.names") %>%
  group_by(multiplier, ptm, replicate) %>%
  summarize(pseudobulk = log2(sum(2^value, na.rm = TRUE)),
            bulk = log2(mean(2^value, na.rm = TRUE)))

df.pb <- data.frame(pseudobulk = subset(df, multiplier == 1)$pseudobulk,
                    bulk = subset(df, multiplier == 100)$bulk,
                    replicate = subset(df, multiplier == 1)$replicate,
                    ptm = subset(df, multiplier == 1)$ptm)

pdf("MS2.log2.sc_vs_bulk_replicate.pdf",
    width = 7,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk)) +
  geom_point() +
  stat_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  facet_wrap(.~as.factor(replicate)) +
  labs(x = "Bulk\nLog2 Mean MS2 Area",
       y = "Pseudobulk\nLog2 Summed MS2 Area",
       title = "Replicate #")
dev.off()
```

## Relative Abundance

### MS1

#### Correlation
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

# correlation all replicates
# pseudobulk here is the mean instead of the sum of single cells
df <- assay(obj.filter.ratio.multi[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj)), by.x = "variable", by.y = "row.names") %>%
  group_by(multiplier, ptm) %>%
  summarize(pseudobulk = mean(value, na.rm = TRUE),
            bulk = mean(value, na.rm = TRUE))
  
df.pb <- data.frame(pseudobulk = subset(df, multiplier == 1)$pseudobulk,
                    bulk = subset(df, multiplier == 100)$bulk,
                    row.names = subset(df, multiplier == 1)$ptm)

pdf("MS1.ratio.log1p.batchCorrect.sc_vs_bulk.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk)) +
  geom_point() +
  stat_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  labs(x = "Bulk\nMean Relative abundance",
       y = "Pseudobulk\nMean Relative abundance")
dev.off()

# correlation by replicate
# pseudobulk here is the mean instead of the sum of single cells
df <- assay(obj.filter.ratio.multi[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
    as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj)), by.x = "variable", by.y = "row.names") %>%
  group_by(multiplier, ptm, replicate) %>%
  summarize(pseudobulk = mean(value, na.rm = TRUE),
            bulk = mean(value, na.rm = TRUE))

df.pb <- data.frame(pseudobulk = subset(df, multiplier == 1)$pseudobulk,
                    bulk = subset(df, multiplier == 100)$bulk,
                    replicate = subset(df, multiplier == 1)$replicate,
                    ptm = subset(df, multiplier == 1)$ptm)

pdf("MS1.ratio.log1p.batchCorrect.sc_vs_bulk_replicate_nocor.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk, color = as.factor(replicate))) +
  geom_point() +  # Adds points
  stat_smooth(method = "lm") +  # Adds a linear model fit
  scale_color_discrete(name = "Replicate") +  # Names the color legend as "Replicate"
  labs(x = "Bulk\nMean Relative abundance",
       y = "Pseudobulk\nMean Relative abundance") +
  theme_Publication(base_size = 16) +  # Custom theme
  theme(legend.position = "top") +
  guides(color = guide_legend(override.aes = list(shape = 19)))
dev.off()

pdf("MS1.ratio.log1p.batchCorrect.sc_vs_bulk_replicate_cor.pdf",
    width = 5,
    height = 5)
ggplot(df.pb, aes(x = bulk, y = pseudobulk, color = as.factor(replicate))) +
  geom_point() +  # Adds points
  stat_smooth(method = "lm") +  # Adds a linear model fit
  stat_cor() +
  scale_color_discrete(name = "Replicate") +  # Names the color legend as "Replicate"
  labs(x = expression(bold(atop("Bulk", Log[2] * " mean relative abundance"))),
       y = expression(bold(atop("Pseudobulk", Log[2] * " mean relative abundance")))) +
  theme_Publication(base_size = 16) +  # Custom theme
  theme(legend.position = "top") +
  guides(color = guide_legend(override.aes = list(shape = 19)))
dev.off()
```

#### Fold change
```{r}
setwd("/Users/ronaldcutler/EinsteinMed\ Dropbox/Ronald\ Cutler/Sidoli-lab/Projects/Single-cell-histone/Analysis/5-sc-vs-bulk")

# calculate fold change of mean bulk vs single cells
df.bulk <- assay(obj.filter.ratio.multi[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj)), by.x = "variable", by.y = "row.names") %>%
  subset(multiplier == 100) %>%
  group_by(ptm) %>%
  summarize(mean = mean(value, na.rm = TRUE))

obj.sc <- obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi[["ptm.ratio"]]$multiplier == 1]
df.sc <- assay(obj.sc, "MS1.ratio.log1p.batchCorrect")
df.sc <- df.sc[match(rownames(df.sc), df.bulk$ptm),]
fc <- df.sc -  df.bulk$mean

# histogram
data <- data.frame(fc = c(as.matrix(fc)))

pdf("MS1.ratio_sc_vs_bulk_histogram.pdf",
    width = 5,
    height = 5)
ggplot(data, aes(x=fc)) + 
  geom_histogram(color="black", fill="white") +
  geom_vline(aes(xintercept = mean(data$fc, na.rm = TRUE)), color = "red", linetype = "dashed", size = 1) +
  annotate("text", x = mean(data$fc, na.rm = TRUE), y = 220, label = "Mean", vjust = 1, hjust = -0.5, color = "red") +
  theme_Publication() +
  labs(y = "Frequency",
       x = "Log2(Single Cell / Bulk) Relative Abundance")
dev.off()

# heatmap
replicate_colors <- c("1" = "#F8766D", "2" = "#00BA38", "3" = "#619CFF")
column_ha = HeatmapAnnotation(Batch = as.factor(obj.sc$replicate),
                              col = list(Batch = replicate_colors))
row_ha = rowAnnotation(Charge = anno_barplot(rowData(obj.sc)$Precursor.Charge))

pdf("MS1.ratio_sc_vs_bulk_heatmap.pdf",
    width = 10,
    height = 10)
Heatmap(fc,
        top_annotation = column_ha,
        right_annotation = row_ha,
        heatmap_legend_param = list(title = "Log2(Single-cell / Bulk mean)"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        row_names_gp = gpar(fontsize = 10))
dev.off()
```

#### H3 single PTM ratios

##### Bulk
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

bulk <- assay(obj.filter.ratio.multi[["sptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  select(starts_with("100")) %>%
  expm1() %>% # inverse of log1p
  rownames_to_column("ptm") %>%
  filter(str_detect(ptm, "H31")) %>%
  filter(!str_detect(ptm, "[un]")) %>%
  filter(!str_detect(ptm, "[hib]")) %>%
  mutate(pos = as.numeric(str_extract(ptm, "(?<=-K)\\d+")),
         aa_pos = factor(paste0("K", str_extract(ptm, "(?<=-K)\\d+")), levels = c("K4", "K9", "K14", "K18", "K23", "K27", "K36")),
         mod = str_extract(ptm, "(?<=\\[)[a-z0-9]+(?=\\])")) %>%
  arrange(pos, mod) %>%
  rowwise() %>%
  mutate(
    Mean = mean(c_across(contains("Slot")), na.rm = FALSE),
    SD = sd(c_across(contains("Slot")), na.rm = FALSE)
  ) %>%
  ungroup()

# calculate cumulative sum for error bars
bulk_cumulative <- bulk %>%
  mutate(mod = factor(mod, levels = c("me", "me2", "me3", "ac"))) %>%
  group_by(pos) %>%
  arrange(pos, desc(mod)) %>%
  mutate(cumulative_mean = cumsum(Mean)) %>%
  ungroup() %>%
  mutate(ymin = cumulative_mean - SD,
         ymax = cumulative_mean + SD)

pdf("MS1.ratio_H3_sptm_100cell.pdf",
    width = 5,
    height = 5)
ggplot(bulk_cumulative, aes(x = aa_pos, y = Mean, fill = mod)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), position = "identity", width = 0.2) +
  labs(x = "Amino Acid Position", y = "Relative Abundance", fill = "Modification") +
  theme_Publication() +
  scale_fill_Publication() +
  theme(legend.position = "top",
        legend.text = element_text(size = 12))
dev.off()

# Negate y-values for downward bars
bulk_cumulative$Mean.flip <- -bulk_cumulative$Mean
bulk_cumulative$ymin.flip <- -bulk_cumulative$ymin
bulk_cumulative$ymax.flip <- -bulk_cumulative$ymax

pdf("MS1.ratio_H3_sptm_100cell_flipped.pdf",
    width = 5,
    height = 5)
ggplot(bulk_cumulative, aes(x = aa_pos, y = Mean.flip, fill = mod)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_errorbar(aes(ymin = ymin.flip, ymax = ymax.flip), position = "identity", width = 0.2) +
  labs(x = "Amino Acid Position", y = "Relative Abundance", fill = "Modification") +
  theme_Publication() +
  scale_fill_Publication() +
  scale_x_discrete(position = "top") + # Move axis to the top
  scale_y_continuous(labels = abs) +
  theme(legend.position = "top",
        legend.text = element_text(size = 12))
dev.off()
```

##### Single cell
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

sc <- assay(obj.filter.ratio.multi[["sptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  select(!starts_with("100")) %>%
  expm1() %>% # inverse of log1p
  rownames_to_column("ptm") %>%
  filter(str_detect(ptm, "H31")) %>%
  filter(!str_detect(ptm, "[un]")) %>%
  filter(!str_detect(ptm, "[hib]")) %>%
  mutate(pos = as.numeric(str_extract(ptm, "(?<=-K)\\d+")),
         aa_pos = factor(paste0("K", str_extract(ptm, "(?<=-K)\\d+")), levels = c("K4", "K9", "K14", "K18", "K23", "K27", "K36")),
         mod = str_extract(ptm, "(?<=\\[)[a-z0-9]+(?=\\])")) %>%
  arrange(pos, mod) %>%
  rowwise() %>%
  mutate(
    Mean = mean(c_across(contains("Slot")), na.rm = FALSE),
    SD = sd(c_across(contains("Slot")), na.rm = FALSE)
  ) %>%
  ungroup()

# calculate cumulative sum for error bars
sc_cumulative <- sc %>%
  mutate(mod = factor(mod, levels = c("me", "me2", "me3", "ac"))) %>%
  group_by(pos) %>%
  arrange(pos, desc(mod)) %>%
  mutate(cumulative_mean = cumsum(Mean)) %>%
  ungroup() %>%
  mutate(ymin = cumulative_mean - SD,
         ymax = cumulative_mean + SD)


pdf("MS1.ratio_H3_sptm_1cell.pdf",
    width = 5,
    height = 5)
ggplot(sc_cumulative, aes(x = aa_pos, y = Mean, fill = mod)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), position = "identity", width = 0.2) +
  labs(x = "H3 Amino Acid Position", y = "Relative Abundance", fill = "Modification") +
  theme_Publication() +
  scale_fill_Publication() +
  theme(legend.position = "top",
        legend.text = element_text(size = 12))
dev.off()
```

### Bulk from extracted histones samples

#### 230816_nabut_bulk_test
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

df <- nabut_bulk_230816 %>% 
  select(contains("0_mmol")) %>%
  rownames_to_column("ptm") %>%
  filter(str_detect(ptm, "H31")) %>%
  filter(!str_detect(ptm, "[un]")) %>%
  filter(!str_detect(ptm, "[hib]")) %>%
  mutate(pos = as.numeric(str_extract(ptm, "(?<=-K)\\d+")),
         aa_pos = factor(paste0("K", str_extract(ptm, "(?<=-K)\\d+")), levels = c("K4", "K9", "K14", "K18", "K23", "K27", "K36")),
         mod = str_extract(ptm, "(?<=\\[)[a-z0-9]+(?=\\])")) %>%
  arrange(pos, mod) %>%
  filter(!is.na(aa_pos)) %>%
  rowwise() %>%
  mutate(
    Mean = mean(c_across(contains("0_mmol")), na.rm = FALSE),
    SD = sd(c_across(contains("0_mmol")), na.rm = FALSE)
  ) %>%
  ungroup() %>%
  filter(Mean != 0)

# calculate cumulative sum for error bars
df_cumulative <- df %>%
  mutate(mod = factor(mod, levels = c("me", "me2", "me3", "ac"))) %>%
  group_by(pos) %>%
  arrange(pos, desc(mod)) %>%
  mutate(cumulative_mean = cumsum(Mean)) %>%
  ungroup() %>%
  mutate(ymin = cumulative_mean - SD,
         ymax = cumulative_mean + SD)


pdf("MS1.ratio_H3_sptm_230816_nabut_bulk_test.pdf",
    width = 5,
    height = 5)
ggplot(df_cumulative, aes(x = aa_pos, y = Mean, fill = mod)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), position = "identity", width = 0.2) +
  labs(x = "Amino Acid Position", y = "Relative Abundance", fill = "Modification") +
  theme_Publication() +
  scale_fill_Publication() +
  theme(legend.position = "top",
        legend.text = element_text(size = 12))
dev.off()
```

#### 231012_broad
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

df <- nabut_bulk_231012 %>% 
  select(contains("auto")) %>%
  rownames_to_column("ptm") %>%
  filter(!str_detect(ptm, "H4")) %>%
  filter(!str_detect(ptm, "H33")) %>%
  filter(!str_detect(ptm, "[un]")) %>%
  filter(!str_detect(ptm, "[hib]")) %>%
  mutate(pos = as.numeric(str_extract(ptm, "(?<=-K)\\d+")),
         aa_pos = factor(paste0("K", str_extract(ptm, "(?<=-K)\\d+")), levels = c("K4", "K9", "K14", "K18", "K23", "K27", "K36", "K79")),
         mod = str_extract(ptm, "(?<=\\[)[a-z0-9]+(?=\\])")) %>%
  arrange(pos, mod) %>%
  filter(!is.na(aa_pos)) %>%
  rowwise() %>%
  mutate(
    Mean = mean(c_across(contains("auto")), na.rm = FALSE),
    SD = sd(c_across(contains("auto")), na.rm = FALSE)
  ) %>%
  ungroup() %>%
  filter(Mean != 0)

# calculate cumulative sum for error bars
df_cumulative <- df %>%
  mutate(mod = factor(mod, levels = c("me1", "me2", "me3", "ac"))) %>%
  group_by(pos) %>%
  arrange(pos, desc(mod)) %>%
  mutate(cumulative_mean = cumsum(Mean)) %>%
  ungroup() %>%
  mutate(ymin = cumulative_mean - SD,
         ymax = cumulative_mean + SD)


pdf("MS1.ratio_H3_sptm_231012_broad.pdf",
    width = 5,
    height = 5)
ggplot(df_cumulative, aes(x = aa_pos, y = Mean, fill = mod)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), position = "identity", width = 0.2) +
  labs(x = "Amino Acid Position", y = "Relative Abundance", fill = "Modification") +
  theme_Publication() +
  scale_fill_Publication() +
  theme(legend.position = "top",
        legend.text = element_text(size = 12))
dev.off()
```

#### Bulk from other studies

##### Bottom-up vs middle-down (Anal Chem 2015)
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

df <- bottomup.sptm %>% 
  rownames_to_column("ptm") %>%
  filter(str_detect(ptm, "H31")) %>%
  filter(!str_detect(ptm, "[un]")) %>%
  filter(!str_detect(ptm, "[hib]")) %>%
  mutate(pos = as.numeric(str_extract(ptm, "(?<=-K)\\d+")),
         aa_pos = factor(paste0("K", str_extract(ptm, "(?<=-K)\\d+")), levels = c("K4", "K9", "K14", "K18", "K23", "K27", "K36")),
         mod = str_extract(ptm, "(?<=\\[)[a-z0-9]+(?=\\])")) %>%
  arrange(pos, mod) %>%
  filter(!is.na(aa_pos)) %>%
  rowwise() %>%
  mutate(
    Mean = mean(c_across(contains("BU")), na.rm = FALSE),
    SD = sd(c_across(contains("BU")), na.rm = FALSE)
  ) %>%
  ungroup() %>%
  filter(Mean != 0)

# calculate cumulative sum for error bars
df_cumulative <- df %>%
  mutate(mod = factor(mod, levels = c("me", "me2", "me3", "ac"))) %>%
  group_by(pos) %>%
  arrange(pos, desc(mod)) %>%
  mutate(cumulative_mean = cumsum(Mean)) %>%
  ungroup() %>%
  mutate(ymin = cumulative_mean - SD,
         ymax = cumulative_mean + SD)


pdf("MS1.ratio_H3_sptm_bottomup-vs_middledown.pdf",
    width = 5,
    height = 5)
ggplot(df_cumulative, aes(x = aa_pos, y = Mean, fill = mod)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), position = "identity", width = 0.2) +
  labs(x = "Amino Acid Position", y = "Relative Abundance", fill = "Modification") +
  theme_Publication() +
  scale_fill_Publication() +
  theme(legend.position = "top",
        legend.text = element_text(size = 12))
dev.off()
```

##### DIA low resolution (Anal Chem 2015)
```{r}
setwd("~/single-cell-histone-ptm/Analysis/5-sc-vs-bulk")

df <- dialowres.sptm %>% 
  rownames_to_column("ptm") %>%
  filter(str_detect(ptm, "H31")) %>%
  filter(!str_detect(ptm, "[un]")) %>%
  filter(!str_detect(ptm, "[hib]")) %>%
  mutate(pos = as.numeric(str_extract(ptm, "(?<=-K)\\d+")),
         aa_pos = factor(paste0("K", str_extract(ptm, "(?<=-K)\\d+")), levels = c("K4", "K9", "K14", "K18", "K23", "K27", "K36")),
         mod = str_extract(ptm, "(?<=\\[)[a-z0-9]+(?=\\])")) %>%
  filter(!is.na(aa_pos)) %>%
  arrange(pos, mod) %>%
  rowwise() %>%
  mutate(
    Mean = mean(c_across(contains("hela")), na.rm = FALSE),
    SD = sd(c_across(contains("hela")), na.rm = FALSE)
  ) %>%
  ungroup() %>%
  filter(Mean != 0)

# calculate cumulative sum for error bars
df_cumulative <- df %>%
  mutate(mod = factor(mod, levels = c("me", "me2", "me3", "ac"))) %>%
  group_by(pos) %>%
  arrange(pos, desc(mod)) %>%
  mutate(cumulative_mean = cumsum(Mean)) %>%
  ungroup() %>%
  mutate(ymin = cumulative_mean - SD,
         ymax = cumulative_mean + SD)


pdf("MS1.ratio_H3_sptm_dialowres.pdf",
    width = 5,
    height = 5)
ggplot(df_cumulative, aes(x = aa_pos, y = Mean, fill = mod)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), position = "identity", width = 0.2) +
  labs(x = "Amino Acid Position", y = "Relative Abundance", fill = "Modification") +
  theme_Publication() +
  scale_fill_Publication() +
  theme(legend.position = "top",
        legend.text = element_text(size = 12))
dev.off()
```