---
title: "8-differential-abundance: nabut"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Libraries
```{r}
library(MultiAssayExperiment)
library(SummarizedExperiment)
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(pheatmap)
library(ComplexHeatmap)
library(ggpubr)
library(reshape2)
library(PCAtools)
library(umap)
library(class)
library(cluster)
library(viridis)
library(boot)
library(EnhancedVolcano)
library(RColorBrewer)
```

## Functions
```{r}
source("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/functions.R")
```

## Load normalized data
- This now has the mixed cells classified into the control or treatment groups
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/7-cell-classification/auto+nabut")

obj.filter.ratio.multi <- readRDS("obj.filter.ratio.multi.RDS")

# add unique replicate ids
obj.filter.ratio.multi$replicate.unique <- factor(as.numeric(factor(obj.filter.ratio.multi$experiment_replicate)))
obj.filter.ratio.multi[["ptm.ratio"]]$replicate.unique <- factor(as.numeric(factor(obj.filter.ratio.multi$experiment_replicate)))
```

# Bulk

## Differential abundance
- Improvement: Calculate confidence interval for log2fc, use glm to calculate fold changes while accounting for batch variables

### All PTMs
```{r}
dir.create("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/bulk")
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/bulk")

res <- differential_abundance(obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 100], 
                                 assay = "MS1.ratio.log1p.batchCorrect",
                                 test_type = "t.test",
                                 group_column = "condition", 
                                 group1 = "control",
                                 group2 = "treat")
res$PTM.Name.Readable <- rowData(obj.filter.ratio.multi[["ptm.ratio"]])$PTM.Name.Readable[match(rownames(res), rownames(rowData(obj.filter.ratio.multi[["ptm.ratio"]])))]

write.csv(res, "nabut_vs_control_bulk.csv")
 
# volcano
# res$volcano.label <- ""
# res$volcano.score <- abs(res$log2fc) * -log10(res$p.value)
# res <- res[order(res$volcano.score, decreasing = TRUE),]
# res$volcano.label[1:20] <- res$ptm[1:20]
# res$direction <- ifelse(res$log2fc > 0, "up", "down")

pdf("nabut_vs_control_volcano_bulk.pdf",
    height = 6,
    width = 6)
EnhancedVolcano(res,
                lab = res$PTM.Name.Readable,
                title = NULL,
                subtitle = NULL,
                caption = NULL,
                x = 'log2FC',
                y = 'p.value',
                pCutoffCol = "p.adj",
                pointSize = 4.0,
                colAlpha = 0.75,
                legendPosition = "none",
                FCcutoff = 0.5,
                pCutoff = 0.05,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = FALSE,
                drawConnectors = FALSE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                arrowheads = FALSE,
                max.overlaps = Inf,
                gridlines.major	= FALSE,
                gridlines.minor = FALSE,
                ylim = c(0, 7),
                xlim = c(-2, 4))
dev.off()

# heatmap sig hits
ptm.ratio.sig <- assay(obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 100], "MS1.ratio.log1p.batchCorrect")[rownames(res[res$p.adj < 0.05,]),]
rownames(ptm.ratio.sig) <- rowData(obj.filter.ratio.multi[["ptm.ratio"]])$PTM.Name.Readable[match(rownames(ptm.ratio.sig),
                                                                                              rownames(rowData(obj.filter.ratio.multi[["ptm.ratio"]])))]

# remove unmodified
ptm.ratio.sig <- ptm.ratio.sig[grep("ac|me", rownames(ptm.ratio.sig)), ]

column_ha = HeatmapAnnotation("Batch" = obj.filter.ratio.multi[,obj.filter.ratio.multi$multiplier == 100]$replicate.unique,
                              "Condition" = obj.filter.ratio.multi[,obj.filter.ratio.multi$multiplier == 100]$condition.predict,
                              col = list("Condition" = c("control" = "#317ec2",
                                                       "treat" = "#c03830"),
                                           "Batch" = c("1" = brewer.pal(5, "Set1")[1],
                                                        "2" = brewer.pal(5, "Set1")[2],
                                                        "3" = brewer.pal(5, "Set1")[3],
                                                        "4" = brewer.pal(5, "Set1")[4],
                                                        "5" = brewer.pal(5, "Set1")[5]
                                                      )
                                        ),
                              annotation_legend_param = list("Condition" = list(direction = "horizontal"),
                                                             "Batch" = list(direction = "horizontal"))
                              )

histone_proteins <- sub("(.*?)K.*", "\\1", rownames(ptm.ratio.sig))


hm <- Heatmap(t(scale(t(ptm.ratio.sig))),
        top_annotation = column_ha,
        column_title = NULL,
        heatmap_legend_param = list(
                  title = "Z-score",
                  direction = "vertical"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        column_split = obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 100]$condition.predict,
        split = histone_proteins,
        cluster_column_slices = FALSE,
        cluster_row_slices = FALSE,
        row_names_gp = gpar(fontsize = 10))

pdf("nabut_vs_control_sig-heatmap_bulk.pdf",
    width = 6,
    height = 4)
draw(hm,
     heatmap_legend_side = "right",
     annotation_legend_side = "right")
dev.off() 
```

### Global PTMs
```{r}
df <- assay(obj.filter.ratio.multi[["global.ratio"]][,obj.filter.ratio.multi$multiplier == 100], "MS1.ratio.log1p.batchCorrect") %>%
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "sample",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj.filter.ratio.multi)), by.x = "sample", by.y = "row.names")

# facet labels
names <- c("Acetylation", "Mono-Methylation", "Di-Methylation", "Tri-Methylation")
codes <- c("ac", "me", "me2", "me3")
facet_labels <- setNames(names, codes)


pdf("nabut_vs_control_global-ptm_boxplot.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = condition.predict,  y = value, fill = condition.predict)) +
  geom_boxplot() +
  theme_Publication() +
  labs(y = "PTM Relative abundance") +
  scale_fill_manual(name = "Condition", 
                    labels = c("Control", "Treat"),
                    values = c("#317ec2" ,"#c03830")) +
  facet_wrap(. ~ ptm, scales = "free_y", nrow = 2, 
             labeller = labeller(ptm = as_labeller(facet_labels))) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  stat_compare_means(aes(group = condition.predict), 
                            method = "t.test", # Choose the appropriate test
                            label = "p.signif",
                            label.x.npc = "center",
                            label.y.npc = "top",
                            hide.ns = TRUE,
                            vjust = 1.5)
dev.off()
```

### H4 Acetylation
```{r}
df <- assay(obj.filter.ratio.multi[["h4.ac.ratio"]][,obj.filter.ratio.multi$multiplier == 100], "MS1.ratio.log1p.batchCorrect") %>%
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "sample",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj.filter.ratio.multi)), by.x = "sample", by.y = "row.names")

# facet labels
names <- c("Unmodified", "Mono-Acetylation", "Di-Acetylation", "Tri-Acetylation", "Tetra-Acetylation")
codes <- c("0-ac", "Mono-ac", "Di-ac", "Tri-ac", "Tetra-ac")
facet_labels <- setNames(names, codes)


pdf("nabut_vs_control_h4.ac_boxplot.pdf",
    width = 4,
    height = 5)
ggplot(df, aes(x = condition,  y = value, fill = condition)) +
  geom_boxplot() +
  theme_Publication() +
  labs(y = "PTM Relative abundance") +
  scale_fill_manual(name = "Condition", 
                    labels = c("Control", "Treat"),
                    values = c("#317ec2" ,"#c03830")) +
  facet_wrap(. ~ ptm, scales = "free_y", nrow = 3, 
             labeller = labeller(ptm = as_labeller(facet_labels))) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        strip.text = element_text(size = 9)) +
  scale_y_continuous(labels = function(x) paste0(x*100, "%")) +
  stat_compare_means(aes(group = condition), 
                            method = "t.test", # Choose the appropriate test
                            label = "p.signif",
                            label.x.npc = "center",
                            label.y.npc = "top",
                            hide.ns = TRUE,
                            vjust = 1.5)
dev.off()
```


# Single Cells

## Differential abundance

### All PTMs

- Improvement: Calculate confidence interval for log2fc
```{r}
dir.create("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

# add predicted cell labels
obj.filter.ratio.multi[["ptm.ratio"]]$condition.predict <- obj.filter.ratio.multi$condition.predict

res.sc <- differential_abundance(obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 1], 
                                 assay = "MS1.ratio.log1p.batchCorrect",
                                 test_type = "wilcox.test",
                                 group_column = "condition.predict", 
                                 group1 = "control",
                                 group2 = "treat")
res.sc$PTM.Name.Readable <- rowData(obj.filter.ratio.multi[["ptm.ratio"]])$PTM.Name.Readable[match(rownames(res.sc), rownames(rowData(obj.filter.ratio.multi[["ptm.ratio"]])))]

write.csv(res.sc, "nabut_vs_control_sc.csv")
 
# volcano
# res$volcano.label <- ""
# res$volcano.score <- abs(res$log2fc) * -log10(res$p.value)
# res <- res[order(res$volcano.score, decreasing = TRUE),]
# res$volcano.label[1:20] <- res$ptm[1:20]
# res$direction <- ifelse(res$log2fc > 0, "up", "down")

pdf("nabut_vs_control_volcano_sc.pdf",
    height = 6,
    width = 6)
EnhancedVolcano(res.sc,
                lab = res.sc$PTM.Name.Readable,
                title = NULL,
                subtitle = NULL,
                caption = NULL,
                x = 'log2FC',
                y = 'p.value',
                pCutoffCol = "p.adj",
                pointSize = 4.0,
                colAlpha = 0.75,
                legendPosition = "none",
                FCcutoff = 0.5,
                pCutoff = 0.05,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = FALSE,
                drawConnectors = TRUE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                arrowheads = FALSE,
                max.overlaps = Inf,
                gridlines.major	= FALSE,
                gridlines.minor = FALSE,
                ylim = c(0, 17),
                xlim = c(-2, 3))
dev.off()

# heatmap sig hits
ptm.ratio.sig <- assay(obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 1], "MS1.ratio.log1p.batchCorrect")[rownames(res.sc[res.sc$p.adj < 0.05,]),]
rownames(ptm.ratio.sig) <- rowData(obj.filter.ratio.multi[["ptm.ratio"]])$PTM.Name.Readable[match(rownames(ptm.ratio.sig),
                                                                                                  rownames(rowData(obj.filter.ratio.multi[["ptm.ratio"]])))]

# remove unmodified
ptm.ratio.sig <- ptm.ratio.sig[grep("ac|me", rownames(ptm.ratio.sig)), ]

column_ha = HeatmapAnnotation("Replicate" = obj.filter.ratio.multi[,obj.filter.ratio.multi$multiplier == 1]$replicate.unique,
                              "Condition" = obj.filter.ratio.multi[,obj.filter.ratio.multi$multiplier == 1]$condition.predict,
                              col = list("Condition" = c("control" = "#317ec2",
                                                       "treat" = "#c03830"),
                                           "Replicate" = c("1" = brewer.pal(5, "Set1")[1],
                                                        "2" = brewer.pal(5, "Set1")[2],
                                                        "3" = brewer.pal(5, "Set1")[3],
                                                        "4" = brewer.pal(5, "Set1")[4],
                                                        "5" = brewer.pal(5, "Set1")[5]
                                                      )
                                        ),
                              annotation_legend_param = list("Condition" = list(direction = "horizontal"),
                                                             "Replicate" = list(direction = "horizontal"))
                              )

histone_proteins <- sub("(.*?)K.*", "\\1", rownames(ptm.ratio.sig))


hm <- Heatmap(t(scale(t(ptm.ratio.sig))),
        top_annotation = column_ha,
        column_title = NULL,
        heatmap_legend_param = list(
                  title = "Z-score",
                  direction = "vertical"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        column_split = obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 1]$condition.predict,
        split = histone_proteins,
        cluster_column_slices = FALSE,
        cluster_row_slices = FALSE,
        row_names_gp = gpar(fontsize = 10))

pdf("nabut_vs_control_sig-heatmap.pdf",
    width = 5,
    height = 7.5)
draw(hm,
     heatmap_legend_side = "top",
     annotation_legend_side = "top")
dev.off() 

# legend
# lgd <- Legend(col_fun = col_fun, title = "legend2", at = c(0, 0.25, 0.5, 0.75, 1),
#     direction = "horizontal")
# 
# pdf("nabut_vs_control_sig-heatmap.pdf",
#     width = 7,
#     height = 7)
# draw(lgd)
# dev.off() 
```

#### Subclustering
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

# Scale the data
ptm.ratio.sig.treat <- assay(obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 1 & obj.filter.ratio.multi$condition.predict == "treat"], "MS1.ratio.log1p.batchCorrect")[rownames(res.sc[res.sc$p.adj < 0.05,]),]

# Scale the data
scaled_data <- t(scale(t(ptm.ratio.sig.treat)))

# Hierarchical clustering on rows
row_dist <- dist(scaled_data) # Compute distance matrix for rows
row_hclust <- hclust(row_dist) # Perform hierarchical clustering

# Hierarchical clustering on columns
col_dist <- dist(t(scaled_data)) # Compute distance matrix for columns
col_hclust <- hclust(col_dist) # Perform hierarchical clustering

# Get column cluster assignments
col_clusters <- cutree(col_hclust, k = 3)

# rename and factorize
# remove outlier cell that is col_clusters == 3
col_clusters[col_clusters == 2] <- "A"
col_clusters[col_clusters == 1] <- "B"
col_clusters_factor <- factor(col_clusters)

# add to multi assay obj
df <- colData(obj.filter.ratio.multi) %>%
  as.data.frame() %>%
  rownames_to_column("row_names") %>% # Convert rownames to a column for matching
  mutate(
    treat.subcluster = case_when(row_names %in%  names(col_clusters)[col_clusters == "A"] ~ "A", 
                          row_names %in% names(col_clusters)[col_clusters == "B"] ~ "B",
                          TRUE ~ NA_character_))
obj.filter.ratio.multi$treat.subcluster <- df$treat.subcluster
obj.filter.ratio.multi[["ptm.ratio"]]$treat.subcluster <- df$treat.subcluster
```

##### Heatmap validation
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

column_ha = HeatmapAnnotation(
  Subcluster = obj.filter.ratio.multi[,obj.filter.ratio.multi$multiplier == 1 & obj.filter.ratio.multi$condition.predict == "treat"]$treat.subcluster,
  Replicate = obj.filter.ratio.multi[,obj.filter.ratio.multi$multiplier == 1 & obj.filter.ratio.multi$condition.predict == "treat"]$replicate.unique,
                              col = list(
                                 Subcluster =  c("A" = brewer.pal(3, "Dark2")[1],
                                                 "B" = brewer.pal(3, "Dark2")[2]),
                                 Replicate = c("4" = brewer.pal(5, "Set1")[4],
                                               "5" = brewer.pal(5, "Set1")[5]
                                                      )
                                        )
                              )

histone_proteins <- sub("^([^-]+)-.*$", "\\1", rownames(ptm.ratio.sig.treat))

pdf("nabut_vs_control_sig-heatmap_treat_subcluster.pdf",
    width = 6,
    height = 8)
Heatmap(scaled_data,
       top_annotation = column_ha,
       column_title = NULL,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        column_split = col_clusters_factor,  # Use cluster assignments for splitting
        split = histone_proteins,
        cluster_column_slices = FALSE,
        cluster_row_slices = FALSE,
        row_names_gp = gpar(fontsize = 10))
dev.off()
```

##### Marker identification
- Use only significant PTMs when comparing control and treat

###### Subcluster A
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

# subset treat cells
obj.filter.ratio.multi.treat <- obj.filter.ratio.multi[["ptm.ratio"]][,obj.filter.ratio.multi$multiplier == 1 & obj.filter.ratio.multi$condition.predict == "treat" & !is.na(obj.filter.ratio.multi$treat.subcluster)]

res.sc.treat.a <- differential_abundance(obj.filter.ratio.multi.treat[rownames(res.sc[res.sc$p.adj < 0.05,]),], 
                                 assay = "MS1.ratio.log1p.batchCorrect",
                                 test_type = "wilcox.test",
                                 group_column = "treat.subcluster", 
                                 group1 = "B",
                                 group2 = "A")

res.sc.treat.a$PTM.Name.Readable <- rowData(obj.filter.ratio.multi.treat)$PTM.Name.Readable[match(rownames(res.sc.treat.a), rownames(obj.filter.ratio.multi.treat))]

write.csv(res.sc.treat.a, "nabut_vs_control_sc_treat_subcluster_a.csv")
 
# volcano
# res$volcano.label <- ""
# res$volcano.score <- abs(res$log2fc) * -log10(res$p.value)
# res <- res[order(res$volcano.score, decreasing = TRUE),]
# res$volcano.label[1:20] <- res$ptm[1:20]
# res$direction <- ifelse(res$log2fc > 0, "up", "down")

pdf("nabut_vs_control_volcano_treat_subcluster_a.pdf",
    height = 6,
    width = 6)
EnhancedVolcano(res.sc.treat.a,
                lab = res.sc.treat.a$PTM.Name.Readable,
                title = NULL,
                subtitle = NULL,
                caption = NULL,
                x = 'log2FC',
                y = 'p.value',
                pCutoffCol = "p.adj",
                pointSize = 4.0,
                colAlpha = 0.75,
                legendPosition = "none",
                FCcutoff = 0.5,
                pCutoff = 0.1,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = FALSE,
                drawConnectors = FALSE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                arrowheads = FALSE,
                max.overlaps = Inf,
                gridlines.major	= FALSE,
                gridlines.minor = FALSE,
                xlim = c(-2.5, 2.5),
                ylim = c(0, 5))
dev.off()
```

###### Subcluster B
- results should be the same as for A but reversed
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

res.sc.treat.b <- differential_abundance(obj.filter.ratio.multi.treat[rownames(res.sc[res.sc$p.adj < 0.05,]),], 
                                 assay = "MS1.ratio.log1p.batchCorrect",
                                 test_type = "wilcox.test",
                                 group_column = "treat.subcluster", 
                                 group1 = "A",
                                 group2 = "B")

write.csv(res.sc.treat.b, "nabut_vs_control_sc_treat_subcluster_b.csv")
 
# volcano
# res$volcano.label <- ""
# res$volcano.score <- abs(res$log2fc) * -log10(res$p.value)
# res <- res[order(res$volcano.score, decreasing = TRUE),]
# res$volcano.label[1:20] <- res$ptm[1:20]
# res$direction <- ifelse(res$log2fc > 0, "up", "down")

pdf("nabut_vs_control_volcano_treat_subcluster_b.pdf",
    height = 6,
    width = 6)
EnhancedVolcano(res.sc.treat.b,
                lab = rownames(res.sc.treat.b),
                title = NULL,
                subtitle = NULL,
                caption = NULL,
                x = 'log2FC',
                y = 'p.value',
                pCutoffCol = "p.adj",
                pointSize = 4.0,
                colAlpha = 0.75,
                legendPosition = "none",
                FCcutoff = 0.5,
                pCutoff = 0.1,
                labSize = 3,
                labCol = 'black',
                labFace = 'bold',
                boxedLabels = FALSE,
                drawConnectors = FALSE,
                widthConnectors = 1.0,
                colConnectors = 'black',
                arrowheads = FALSE,
                max.overlaps = Inf,
                gridlines.major	= FALSE,
                gridlines.minor = FALSE)
dev.off()
```

##### Heatmap of Subcluster markers
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

# heatmap sig hits
subcluster.markers.sig <- assay(obj.filter.ratio.multi.treat, "MS1.ratio.log1p.batchCorrect")[unique(c(rownames(res.sc.treat.a[res.sc.treat.a$p.adj < 0.1,]),
                                                                                         rownames(res.sc.treat.b[res.sc.treat.b$p.adj < 0.1,]))),]

rownames(subcluster.markers.sig) <- rowData(obj.filter.ratio.multi.treat)$PTM.Name.Readable[match(rownames(subcluster.markers.sig),
                                                                                                  rownames(rowData(obj.filter.ratio.multi.treat)))]

# remove unmodified
subcluster.markers.sig <- subcluster.markers.sig[grep("ac|me", rownames(subcluster.markers.sig)), ]

column_ha = HeatmapAnnotation(
                              Condition = obj.filter.ratio.multi.treat$condition.predict,
                              Replicate = obj.filter.ratio.multi.treat$replicate.unique,
                              Subcluster = obj.filter.ratio.multi.treat$treat.subcluster,
                              col = list(Condition = c("treat" = "#c03830"),
                                           Replicate = c("4" = brewer.pal(5, "Set1")[4],
                                                        "5" = brewer.pal(5, "Set1")[5]),
                                         Subcluster = c("A" = brewer.pal(3, "Dark2")[1],
                                                        "B" = brewer.pal(3, "Dark2")[2])
                                        )
                              )

histone_proteins <- sub("(.*?)K.*", "\\1", rownames(subcluster.markers.sig))

pdf("nabut_vs_control_sig-heatmap_treat_subcluster_sig-markers.pdf",
    width = 5,
    height = 3.5)
Heatmap(t(scale(t(subcluster.markers.sig))),
        top_annotation = column_ha,
        column_title = NULL,
        heatmap_legend_param = list(title = "Z-score"),
        cluster_rows = TRUE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        column_split = obj.filter.ratio.multi.treat$treat.subcluster,  # Use cluster assignments for splitting
        split = histone_proteins,
        cluster_column_slices = FALSE,
        cluster_row_slices = FALSE,
        row_names_gp = gpar(fontsize = 10))
dev.off() 
```

### Global PTMs
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

df <- assay(obj.filter.ratio.multi[["global.ratio"]][,obj.filter.ratio.multi$multiplier == 1], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "sample",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj.filter.ratio.multi)), by.x = "sample", by.y = "row.names")

# facet labels
names <- c("Acetylation", "Mono-Methylation", "Di-Methylation", "Tri-Methylation")
codes <- c("ac", "me", "me2", "me3")
facet_labels <- setNames(names, codes)


pdf("nabut_vs_control_global-ptm_violin.pdf",
    width = 5,
    height = 5)
ggplot(df, aes(x = condition.predict,  y = value, fill = condition.predict)) +
  geom_violin() +
  geom_boxplot(width = 0.2, fill = "white") +
  theme_Publication() +
  labs(y = "Global Relative Abundance") +
  scale_fill_manual(name = "Condition", 
                    labels = c("Control", "Treat"),
                    values = c("#317ec2" ,"#c03830")) +
  facet_wrap(. ~ ptm, scales = "free_y", nrow = 2, 
             labeller = labeller(ptm = as_labeller(facet_labels))) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  stat_compare_means(aes(group = condition.predict), 
                            method = "wilcox.test", # Choose the appropriate test
                            label = "p.format",
                            label.x.npc = "center",
                            label.y.npc = "top",
                            hide.ns = TRUE,
                            vjust = 1.5,
                            size = 5)
dev.off()
```

### H4 Acetylation
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut/single-cell")

df <- assay(obj.filter.ratio.multi[["h4.ac.ratio"]][,obj.filter.ratio.multi$multiplier == 1], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "sample",
    values_to = "value") %>%
  merge(as.data.frame(colData(obj.filter.ratio.multi)), by.x = "sample", by.y = "row.names") %>%
  mutate(ptm = factor(ptm, levels = c("0-ac", "Mono-ac", "Di-ac", "Tri-ac", "Tetra-ac")))

# facet labels
names <- c("Unmodified", "Mono-Acetylation", "Di-Acetylation", "Tri-Acetylation", "Tetra-Acetylation")
codes <-c("0-ac", "Mono-ac", "Di-ac", "Tri-ac", "Tetra-ac")
facet_labels <- setNames(names, codes)

pdf("nabut_vs_control_h4.ac_violin.pdf",
    width = 5,
    height = 6)
ggplot(df, aes(x = condition.predict,  y = value, fill = condition.predict)) +
  geom_violin() +
  geom_boxplot(width = 0.2, fill = "white") +
  theme_Publication() +
  labs(y = "H4 acetylation Relative Abundance") +
  scale_fill_manual(name = "Condition", 
                    labels = c("Control", "Treat"),
                    values = c("#317ec2" ,"#c03830")) +
  facet_wrap(. ~ ptm, scales = "free_y", nrow = 3, 
             labeller = labeller(ptm = as_labeller(facet_labels))) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        strip.text = element_text(size = 13)) +
  #scale_y_continuous(labels = function(x) paste0(x*100, "%")) +
  stat_compare_means(aes(group = condition.predict), 
                            method = "wilcox.test", # Choose the appropriate test
                            label = "p.format",
                            label.x.npc = "center",
                            label.y.npc = "top",
                            hide.ns = TRUE,
                            vjust = 1.5,
                     size = 5)
dev.off()
```

# Export data and save object
```{r}
setwd("/Users/ronaldcutler/Dropbox\ (EinsteinMed)/Sidoli-lab/Projects/Single-cell-histone/Analysis/8-differential-abundance/auto+nabut")

# export new cluster assignments
write.csv(colData(obj.filter.ratio.multi), "obj.filter.ratio.multi.meta.csv")

# save
saveRDS(obj.filter.ratio.multi, "obj.filter.ratio.multi.RDS")
```

