---
title: "biological variability: auto+nabut"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes
## Variance mean ratio vs coefficient of variation
Variance Mean Ratio (Index of Dispersion)
Interpretation: The variance mean ratio is greater than 1 when data are overdispersed relative to the Poisson distribution (variance exceeds mean), less than 1 for underdispersion (variance is less than the mean), and exactly 1 if the data perfectly follows a Poisson distribution.
Use Case: This metric is particularly insightful for count data or in fields like ecology, genetics, or epidemiology where the data generation process might be expected to follow or deviate from Poisson-like statistics.
Coefficient of Variation (CV)
Interpretation: The CV is a normalized measure of dispersion of a probability distribution and is expressed as a percentage. It’s particularly useful when the mean of the data is not near zero.
Use Case: CV is excellent for comparing the degree of variation between datasets with different units or scales and is commonly used across a broad range of applications, including but not limited to finance, meteorology, and bioinformatics.
Choosing Between Variance Mean Ratio and CV
The choice between these two depends on:

Type of Data: If you are dealing with count data and you want to assess the over or underdispersion relative to a Poisson process, the variance mean ratio is the appropriate choice. If your data are ratio-scaled and the mean is not close to zero, the coefficient of variation might provide more intuitive insights.
Purpose of Analysis: Consider what aspect of the data's variability is most important to communicate:
Use the variance mean ratio if the goal is to understand how the data conform to a Poisson distribution, which is particularly relevant in statistical modeling where assumptions about data distribution impact model choice.
Use the coefficient of variation if the goal is to compare the variability of different datasets on a relative scale, especially when these datasets are not necessarily counts and have significantly different mean values.

# Preparation

## Libraries
```{r}
library(MultiAssayExperiment)
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(pheatmap)
library(ComplexHeatmap)
library(ggpubr)
library(reshape2)
library(PCAtools)
library(umap)
library(class)
library(cluster)
library(viridis)
library(boot)
library(EnhancedVolcano)
library(corrplot)
library(ggpubr)
library(betareg)
library(RColorBrewer)
```

## Functions
```{r}
source("~/single-cell-histone-ptm/functions.R")
```

## Load normalized cell data
- 22 single cells
```{r}
setwd("~/single-cell-histone-ptm/4-normalization/auto")

obj.filter.ratio.multi.cells <- readRDS("obj.filter.ratio.multi.RDS")

## subset single cells
obj.filter.ratio.multi.sc <- obj.filter.ratio.multi.cells[,obj.filter.ratio.multi.cells$multiplier == 1]
ncol(obj.filter.ratio.multi.sc[["ptm.ratio"]])
```

## Load normalized histone standard data
- 12 standards
```{r}
setwd("~/single-cell-histone-ptm/6-technical-variability/auto")

obj.filter.ratio.multi.standards.10pg <- readRDS("obj.filter.ratio.multi.10pg.RDS")

ncol(obj.filter.ratio.multi.standards.10pg)
```

# Analysis

## With technical variation

### Without unmodified peptidoforms
```{r}
# filter out unmodified
peptides <- rownames(obj.filter.ratio.multi.sc)
peptides <- peptides[!(grepl("\\[un\\]", peptides) & !grepl("\\[(me|ac|me2|me3|su|hib)\\]", peptides))]
obj.filter.ratio.multi.sc.no_unmod <- obj.filter.ratio.multi.sc[rownames(obj.filter.ratio.multi.sc) %in% peptides,]

# make the histone ptms readable
rownames(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]]) <- ptm_readable_translation(rownames(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]]))
```

#### Peptidoform level

##### PCA
```{r}
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation")
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation/without_unmodified_peptidoforms")
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation/without_unmodified_peptidoforms/pca")
setwd("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation/without_unmodified_peptidoforms/pca")

p <- pca(t(assay(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect")),
         metadata = rowData(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]]),
         center = TRUE, 
         scale = TRUE)

pdf("MS1.ratio.log1p.batchCorrect_screeplot_peptidoform.pdf",
    width = 5,
    height = 5)
screeplot(p,
          components = getComponents(p)[1:10])
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_eigencor_peptidoform.pdf",
    width = 5,
    height = 5)
eigencorplot(p,
             main = "Principle component correlations",
             cexMain = 1.5,
             metavars = c("Protein.Gene", "Peptide.Sequence", "Precursor.Charge"),
             col = viridis(100),
             colCorval = 'firebrick',
             fontCorval = 2,
             cexCorval = 0.5,
             rotLabX = 45,
             posColKey = 'top')
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_protein_pca_peptidoform.pdf",
    width = 7,
    height = 5)
biplot(p,
       colby = "Protein.Gene", 
       hline = 0, 
       vline = 0,
       lab = NULL) +
  labs(color = "Protein") +
  theme_Publication() +
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.margin=margin())
dev.off()
```

##### Variance
- Positive relationship between variance and mean
- Log2 transform counts are more normally distributed
- polynomial regression when log transform counts
```{r}
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation/without_unmodified_peptidoforms/variance")
setwd("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation/without_unmodified_peptidoforms/variance")

# get relative abundance
df <- assay(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>%
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  group_by(ptm) %>%
  summarize(mean = mean(value, na.rm = TRUE),
            mean.log1p = log1p(mean),
            variance = var(value, na.rm = TRUE),
            vmr = variance/mean) %>%
  merge(as.data.frame(rowData(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]])), by.x = "ptm", by.y = "row.names")


# plot variance
df.plot <- df[order(df$variance, decreasing = TRUE),] # order
df.plot$ptm <- factor(df.plot$ptm,
                      levels = rev(df.plot$ptm))

pdf("MS1.ratio.log1p.batchCorrect_variance_all.pdf",
    width = 6,
    height = 5)
ggplot(df.plot, aes(x=ptm, y=variance, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Variance",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()) +
  ylim(c(0, 0.03))
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_variance_top_20.pdf",
    width = 6,
    height = 5)
ggplot(df.plot[1:20,], aes(x=ptm, y=variance, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Variance",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()) +
  ylim(c(0, 0.03))
dev.off()

# var vs mean
pdf("MS1.ratio.log1p.batchCorrect_mean_vs_variance.pdf",
    width = 7,
    height = 5)
ggplot(df, aes(x = mean, y = variance)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  labs(y = "Variance",
       x = "Mean Relative Abundance",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin())
dev.off()

# plot vmr
df.plot <- df[order(df$vmr, decreasing = TRUE),] # order
df.plot$ptm <- factor(df.plot$ptm,
                      levels = rev(df.plot$ptm))

pdf("MS1.ratio.log1p.batchCorrect_vmr_all.pdf",
    width = 6,
    height = 5)
ggplot(df.plot, aes(x=ptm, y=vmr, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Variance Mean Ratio",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()) +
  ylim(c(0,0.07))
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_vmr_top_20.pdf",
    width = 6,
    height = 5)
ggplot(df.plot[1:20,], aes(x=ptm, y=vmr, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Variance Mean Ratio",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()) +
  ylim(c(0,0.07))
dev.off()

# vmr vs mean
pdf("MS1.ratio.log1p.batchCorrect_mean_vs_vmr.pdf",
    width = 7,
    height = 5)
ggplot(df, aes(x = mean, y = vmr)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  labs(y = "VMR",
       x = "Mean Relative Abundance",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin())
dev.off()

# var vs log2 mean
model <- lm(variance ~ poly(mean.log1p, 2, raw=TRUE), data=df)
coefficients <- coef(model)
equation <- sprintf("y = %.2f + %.2f*x + %.2f*x^2",
                    coefficients[1], coefficients[2], coefficients[3])
pdf("MS1.ratio.log1p.batchCorrect_mean.log1p_vs_variance.pdf",
    width = 7,
    height = 5)
ggplot(df, aes(x = mean.log1p, y = variance)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "blue") +
  stat_cor() +
  theme_Publication() +
  labs(y = "Variance",
       x = "Log2(Mean Relative Abundance)",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin()) +
  annotate("text", x = Inf, y = Inf, label = equation, hjust = 1.5, vjust = 5, size = 4, color = "black")
dev.off()

# vmr vs log2 mean
pdf("MS1.ratio.log1p.batchCorrect_mean.log1p_vs_vmr.pdf",
    width = 7,
    height = 5)
ggplot(df, aes(x = mean.log1p, y = vmr)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  labs(y = "VMR",
       x = "Log2(Mean Relative Abundance)",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin())
dev.off()
```

##### Coefficient of variation
- Inverse relationship between mean and CV^2
```{r}
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation/without_unmodified_peptidoforms/coefficient_of_variation")
setwd("~/single-cell-histone-ptm/7-biological-variability/auto/with_technical_variation/without_unmodified_peptidoforms/coefficient_of_variation")

# get relative abundance
df <- assay(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>%
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  group_by(ptm) %>%
  summarize(mean = mean(value, na.rm = TRUE),
            mean.log1p = log1p(mean),
            cv = sd(value, na.rm = TRUE)/mean,
            cv.squared = cv^2) %>%
  merge(as.data.frame(rowData(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]])), by.x = "ptm", by.y = "row.names")


# plot cv
df.plot <- df[order(df$cv, decreasing = TRUE),] # order
df.plot$ptm <- factor(df.plot$ptm,
                      levels = rev(df.plot$ptm))

pdf("MS1.ratio.log1p.batchCorrect_cv_all.pdf",
    width = 5,
    height = 7)
ggplot(df.plot, aes(x=ptm, y=cv, fill=mean.log1p)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_distiller(palette = 'RdBu', 
                       limits = c(0, 0.75), # Set legend scale from 0 to 1
                       breaks = c(0, 0.25, 0.5, 0.75)
                       oob = scales::squish) + # Squish out-of-bound values to 0-1
  labs(y = "Coefficient of Variation",
       x = "Histone PTM",
       fill = "Mean\nRelative\nAbundance",
       subtitle = "With Technical Noise") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      legend.title = element_text(size = 10)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     breaks = c(seq(0, 1, by = 0.25)),
                     limits = c(0, 1))
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_cv_top_20.pdf",
    width = 5,
    height = 6)
ggplot(df.plot[1:20,], aes(x=ptm, y=cv, fill=mean.log1p)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  #scale_fill_manual(values = protein_colors) +
  scale_fill_distiller(palette = 'RdBu') +
  labs(y = "Coefficient of Variation",
       x = "Histone PTM",
       fill = "Mean\nRelative\nAbundance",
       subtitle = "With Technical Noise") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      legend.title = element_text(size = 10)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     breaks = c(seq(0, 1.25, by = 0.25)),
                     limits = c(0, 1.25))
dev.off()

# heatmap
column_ha = HeatmapAnnotation(Replicate = obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]]$replicate,
                              "Total MS1" = obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]]$total.ms1,
                              col = list(Replicate = c("1" = brewer.pal(3, "Set1")[1],
                                                       "2" = brewer.pal(3, "Set1")[2],
                                                       "3" = brewer.pal(3, "Set1")[3])))

mat <- assay(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect")
mat <- mat[match(df.plot$ptm, rownames(mat)),]

# Scale each row to have mean 0 and standard deviation 1
mat.scaled <- t(scale(t(mat), scale = TRUE, center = FALSE))

row_ha = rowAnnotation(CV = anno_barplot(df.plot$cv))

pdf("MS1.ratio.log1p.batchCorrect_cv_heatmap.pdf",
    width = 10,
    height = 10)
Heatmap(mat.scaled,
        top_annotation = column_ha,
        right_annotation = row_ha,
        heatmap_legend_param = list(title = "Scaled Relative Abundance"),
        cluster_rows = FALSE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# cv vs mean
pdf("MS1.ratio.log1p.batchCorrect_mean_vs_cv.pdf",
    width = 6,
    height = 5)
ggplot(df, aes(x = mean, y = cv)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm", color = "blue") +
  stat_cor() +
  theme_Publication() +
  scale_color_manual(values = protein_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = c(seq(0, 1.5, by = 0.25))) +
  labs(y = "Coefficient of Variation",
       x = "Mean Relative Abundance",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin())
  # annotate("text", x = Inf, y = Inf, label = equation, hjust = 1.5, vjust = 5, size = 4, color = "black")
dev.off()

# cv vs log2 mean
model <- lm(cv.squared ~ poly(mean.log1p, 2, raw=TRUE), data=df)
coefficients <- coef(model)
equation <- sprintf("y = %.2f + %.2f*x + %.2f*x^2",
                    coefficients[1], coefficients[2], coefficients[3])
pdf("MS1.ratio.log1p.batchCorrect_mean.log1p_vs_cv.pdf",
    width = 7,
    height = 5)
ggplot(df, aes(x = mean.log1p, y = cv)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "blue") +
  stat_cor() +
  theme_Publication() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = c(seq(0, 1.5, by = 0.25))) +
  labs(y = "Coefficient of Variation",
       x = "Log2(Mean Relative Abundance)",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin())+
  annotate("text", x = Inf, y = Inf, label = equation, hjust = 1.5, vjust = 5, size = 4, color = "black")
dev.off()

# plot cv2
df.plot <- df[order(df$cv.squared, decreasing = TRUE),] # order
df.plot$ptm <- factor(df.plot$ptm,
                      levels = rev(df.plot$ptm))

pdf("MS1.ratio.log1p.batchCorrect_cv2_all.pdf",
    width = 6,
    height = 5)
ggplot(df.plot, aes(x=ptm, y=cv.squared, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Coefficient of Variation Squared",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank()) +
  ylim(c(0, 1.4))
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_cv2_top_20.pdf",
    width = 6,
    height = 5)
ggplot(df.plot[1:20,], aes(x=ptm, y=cv.squared, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Coefficient of Variation Squared",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank())  +
  ylim(c(0, 1.4))
dev.off()

# cv2 vs mean
model <- lm(cv.squared ~ poly(mean, 2, raw=TRUE), data=df)
coefficients <- coef(model)
equation <- sprintf("y = %.2f + %.2f*x + %.2f*x^2",
                    coefficients[1], coefficients[2], coefficients[3])
pdf("MS1.ratio.log1p.batchCorrect_mean_vs_cv2.pdf",
    width = 7,
    height = 5)
ggplot(df, aes(x = mean, y = cv.squared)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "blue") +
  stat_cor() +
  theme_Publication() +
  labs(y = "Coefficient of Variation Squared",
       x = "Mean Relative Abundance",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin()) +
  annotate("text", x = Inf, y = Inf, label = equation, hjust = 1.5, vjust = 5, size = 4, color = "black")
dev.off()

# cv2 vs log2 mean
model <- lm(cv.squared ~ poly(mean.log1p, 2, raw=TRUE), data=df)
coefficients <- coef(model)
equation <- sprintf("y = %.2f + %.2f*x + %.2f*x^2",
                    coefficients[1], coefficients[2], coefficients[3])
pdf("MS1.ratio.log1p.batchCorrect_mean.log1p_vs_cv2.pdf",
    width = 7,
    height = 5)
ggplot(df, aes(x = mean.log1p, y = cv.squared)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "blue") +
  stat_cor() +
  theme_Publication() +
  labs(y = "Coefficient of Variation Squared",
       x = "Log2(Mean Relative Abundance)",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin()) +
  annotate("text", x = Inf, y = Inf, label = equation, hjust = 1.5, vjust = 5, size = 4, color = "black")
dev.off()
```

## Without technical variation
- See technical variability script for plots of technical variation
- Note that not all ptms in cells were quantified in standards

### Adjust for technical variation
- This is assuming that the total variation is the sum of the technical and biologial variation
- This accounts for the fact that the data is log1p transformed
- Cannot subtract CV from each other as CV is a ratio rather than an additive measure
- If technical is higher than biological, than this ptm is filtered out
- Improvement: adjust for technical variance using a mixed effects model lmer
```{r}
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/without_technical_variation")
setwd("~/single-cell-histone-ptm/7-biological-variability/auto/without_technical_variation")

# get the technical variance
tech.var <- assay(obj.filter.ratio.multi.standards.10pg, "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  group_by(ptm) %>%
  summarize(mean = mean(value, na.rm = TRUE),
            var = sd(value, na.rm = TRUE)^2,
            cv = sd(value, na.rm = TRUE) / mean) %>%
  merge(as.data.frame(rowData(obj.filter.ratio.multi.standards.10pg)), by.x = "ptm", by.y = "row.names")


# get the cell-to-cell variance
bio.var <- assay(obj.filter.ratio.multi.sc[["ptm.ratio"]], "MS1.ratio.log1p.batchCorrect") %>% 
  as.data.frame() %>%
  rownames_to_column("ptm") %>%
  pivot_longer(
    cols = -ptm, # This keeps the 'id' column fixed
    names_to = "variable",
    values_to = "value") %>%
  group_by(ptm) %>%
  summarize(mean = mean(value, na.rm = TRUE),
            var = sd(value, na.rm = TRUE)^2,
            cv = sd(value, na.rm = TRUE) / mean) %>%
  merge(as.data.frame(rowData(obj.filter.ratio.multi.sc[["ptm.ratio"]])), by.x = "ptm", by.y = "row.names")

# merge with bio as ref
# x = bio, y = tech
var.merged <- merge(bio.var, tech.var, by = "ptm")

# get adjusted biological variation 
var.merged <- var.merged %>%
  mutate(var.adj = var.x - var.y, # adjust for technical variance
         cv.adj = sqrt(var.adj) / mean.x) %>% # convert to coefficient of variance
  filter(var.adj > 0) # if there are negative variances, then filter these out

bio.var.adj <- data.frame(ptm = var.merged$ptm,
                          Protein.Gene = var.merged$Protein.Gene,
                          Peptide.Sequence = var.merged$Peptide.Sequence.x,
                          Precursor.Charge = var.merged$Precursor.Charge.x,
                          mean.bio = var.merged$mean.x,
                          var.bio = var.merged$var.x,
                          cv.bio = var.merged$cv.x,
                          mean.tech = var.merged$mean.y,
                          var.tech = var.merged$var.y,
                          cv.tech = var.merged$cv.y,
                          var.bio.adj = var.merged$var.adj,
                          cv.bio.adj = var.merged$cv.adj)

write.csv(bio.var.adj, 
          "MS1.ratio.log1p.CV.adjusted.csv",
          row.names = FALSE)

# pairwise comparison
bio.var.adj.plot <- melt(bio.var.adj[,c("ptm", "cv.tech", "cv.bio")])
pdf("MS1_CV_bio_vs_tech_pairwise.pdf",
    width = 4.5,
    height = 5)
ggpaired(bio.var.adj.plot, x = "variable", y = "value", id = "ptm",
         fill =  c("forestgreen","#D95F02"),
         line.color = "gray",
         line.size = 0.4) +
    stat_compare_means(comparisons = list(c("cv.bio", "cv.tech")),
                      paired = TRUE,
                     method = "wilcox.test",
                     label = "p.format",
                     tip.length = 0,
                     vjust = 0.5,
                     size = 8) +
    labs(x = "Sample type", y = "Coefficient of Variation") +
    scale_x_discrete(labels = c("Histone Standard\n 10 pg", "Single-cells")) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1), breaks = c(seq(0, 1.5, by = 0.25))) +
    theme_Publication()
dev.off()
```

### Without unmodified peptidoforms
```{r}
# filter out unmodified
peptides <- bio.var.adj$ptm
peptides <- peptides[!(grepl("\\[un\\]", peptides) & !grepl("\\[(me|ac|me2|me3|su|hib)\\]", peptides))]
bio.var.adj.no_unmod <- bio.var.adj[bio.var.adj$ptm %in% peptides,]

# make the histone ptms readable
bio.var.adj.no_unmod$ptm <- ptm_readable_translation(bio.var.adj.no_unmod$ptm)
```

##### Variance
- Positive relationship between variance and mean
- Log2 transform counts are more normally distributed
- polynomial regression when log transform counts
```{r}
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/without_technical_variation/without_unmodified_peptidoforms")
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/without_technical_variation/without_unmodified_peptidoforms/variance")
setwd("~/single-cell-histone-ptm/7-biological-variability/auto/without_technical_variation/without_unmodified_peptidoforms/variance")

# plot variance
df.plot <- bio.var.adj.no_unmod[order(bio.var.adj.no_unmod$var.bio.adj, decreasing = TRUE),] # order
df.plot$ptm <- factor(df.plot$ptm,
                      levels = rev(df.plot$ptm))

pdf("MS1.ratio.log1p.batchCorrect_variance_all.pdf",
    width = 6,
    height = 5)
ggplot(df.plot, aes(x=ptm, y=var.bio.adj, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Adjusted Variance",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank())
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_variance_top_20.pdf",
    width = 6,
    height = 5)
ggplot(df.plot[1:20,], aes(x=ptm, y=var.bio.adj, fill=Protein.Gene)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_manual(values = protein_colors) +
  labs(y = "Adjusted Variance",
       x = "Histone PTM",
       fill = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank())
dev.off()

# var vs mean
pdf("MS1.ratio.log1p.batchCorrect_mean_vs_variance.pdf",
    width = 7,
    height = 5)
ggplot(df.plot, aes(x = mean.bio, y = var.bio.adj)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm") +
  stat_cor() +
  theme_Publication() +
  labs(y = "Adjusted Variance",
       x = "Mean Relative Abundance",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin())
dev.off()
```

##### Coefficient of variation
- Inverse relationship between mean and CV^2
```{r}
dir.create("~/single-cell-histone-ptm/7-biological-variability/auto/without_technical_variation/without_unmodified_peptidoforms/coefficient_of_variation")
setwd("~/single-cell-histone-ptm/7-biological-variability/auto/without_technical_variation/without_unmodified_peptidoforms/coefficient_of_variation")

# plot cv
df.plot <- bio.var.adj.no_unmod[order(bio.var.adj.no_unmod$cv.bio.adj, decreasing = TRUE),] # order
df.plot$ptm <- factor(df.plot$ptm,
                      levels = rev(df.plot$ptm))

pdf("MS1.ratio.log1p.batchCorrect_cv_all.pdf",
    width = 5,
    height = 6)
ggplot(df.plot, aes(x=ptm, y=cv.bio.adj, fill=mean.bio)) +
    geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  scale_fill_distiller(palette = 'RdBu') +
  labs(y = "Adjusted Coefficient of Variation",
       x = "Histone PTM",
       fill = "Mean\nRelative\nAbundance",
       subtitle = "After Removal of Technical Noise") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      legend.title = element_text(size = 10)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     breaks = c(seq(0, 1, by = 0.25)),
                     limits = c(0, 1))
dev.off()

pdf("MS1.ratio.log1p.batchCorrect_cv_top_20.pdf",
    width = 5,
    height = 6)
ggplot(df.plot[1:20,], aes(x=ptm, y=cv.bio.adj, fill=mean.bio)) +
  geom_bar(stat="identity", width=0.95, position = position_dodge(width=1)) +
  geom_text(aes(label=ptm), vjust=0.5, hjust = -0.05, angle=0, position=position_dodge(width=1), size = 3) +
  coord_flip() +
  theme_Publication() +
  #scale_fill_manual(values = protein_colors) +
  scale_fill_distiller(palette = 'RdBu') +
  labs(y = "Adjusted Coefficient of Variation",
       x = "Histone PTM",
       fill = "Mean\nRelative\nAbundance",
       subtitle = "After Removal of Technical Noise") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      legend.title = element_text(size = 10)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     breaks = c(seq(0, 1, by = 0.25)),
                     limits = c(0, 1))
dev.off()

# heatmap
obj.filter.ratio.multi.sc.no_unmod.temp <- obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]][rownames(obj.filter.ratio.multi.sc.no_unmod[["ptm.ratio"]]) %in% df.plot$ptm,]

column_ha = HeatmapAnnotation(Replicate = obj.filter.ratio.multi.sc.no_unmod.temp$replicate,
                              "Total Log2(MS1)" = log2(obj.filter.ratio.multi.sc.no_unmod.temp$total.ms1),
                              col = list(Replicate = c("1" = brewer.pal(3, "Set1")[1],
                                                       "2" = brewer.pal(3, "Set1")[2],
                                                       "3" = brewer.pal(3, "Set1")[3])))

mat <- assay(obj.filter.ratio.multi.sc.no_unmod.temp, "MS1.ratio.log1p.batchCorrect")
mat <- mat[match(df.plot$ptm, rownames(mat)),]

# Scale each row to have mean 0 and standard deviation 1
mat.scaled <- t(scale(t(mat), scale = TRUE, center = FALSE))

row_ha = rowAnnotation("Adjusted CV" = anno_barplot(df.plot$cv.bio.adj))

pdf("MS1.ratio.log1p.batchCorrect_cv_heatmap.pdf",
    width = 8,
    height = 7)
Heatmap(mat.scaled,
        top_annotation = column_ha,
        right_annotation = row_ha,
        heatmap_legend_param = list(title = "Scaled Relative Abundance"),
        cluster_rows = FALSE,
        cluster_columns = TRUE,
        show_column_names = FALSE,
        row_names_gp = gpar(fontsize = 10))
dev.off()

# cv vs mean
# model <- lm(cv.bio.adj ~ poly(mean.bio, 2, raw=TRUE), data=df.plot)
# coefficients <- coef(model)
# equation <- sprintf("y = %.2f + %.2f*x + %.2f*x^2",
#                     coefficients[1], coefficients[2], coefficients[3])
pdf("MS1.ratio.log1p.batchCorrect_mean_vs_cv.pdf",
    width = 7,
    height = 5)
ggplot(df.plot, aes(x = mean.bio, y = cv.bio.adj)) +
  geom_point(aes(color = Protein.Gene)) +
  geom_smooth(method = "lm", color = "blue") +
  stat_cor() +
  theme_Publication() +
  scale_color_manual(values = protein_colors) +
  labs(y = "Adjusted Coefficient of Variation",
       x = "Log1p Mean Relative Abundance",
       color = "Protein") +
  theme(legend.position = "right",
      legend.direction = "vertical",
      legend.margin=margin())
dev.off()
```