---
title: "HeLa treated Treatment FACS Single-Cell Analysis"
output: html_document
date: "2025-08-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparation

## Libraries
```{r}
library(tidyr)
library(dplyr)
library(tibble)
library(ggplot2)
library(ggpubr)
library(ComplexHeatmap)
library(viridis)
library(RColorBrewer)
library(rstatix)

library(SummarizedExperiment)
```

## Working Directory
```{r}
working.dir <- "~/EinsteinMed Dropbox/Ronald Cutler/Sidoli-lab/Projects/sc-hPTM/Experiments/250806 FACS"
```

## Functions
```{r}
setwd(working.dir)

source("functions.R")
```

## Load Data
```{r}
setwd(paste(working.dir, "1-data", sep = "/"))

# load conditions
control <- read_flowjo_csv_to_SE("control_comp.csv")

treated <- read_flowjo_csv_to_SE("nabut_comp.csv")

# combine into single se object
control$condition <- "Control"
treated$condition <- "Treated"

## ensure assay order matches, then align features
stopifnot(setequal(assayNames(control), assayNames(treated)))
assays(treated) <- assays(treated)[assayNames(control)]                  # reorder assays to match control
common <- intersect(rownames(control), rownames(treated))            # shared channels/features
controla <- control[common, , drop = FALSE]
treateda <- treated[common, , drop = FALSE]

## combine by columns (adds cells)
se <- SummarizedExperiment::cbind(controla, treateda)
colnames(se) <- make.unique(colnames(se))   # tidy duplicates if any

# annotate antibodies
map <- c("Comp-Alexa Fluor 488-A"="H3K14ac", 
         "Comp-Alexa Fluor 555-A"="H3K23ac", 
         "Comp-APC-A"="H3",
         "Comp-FxCycle Violet-A"="Cell.Cycle",
         "Comp-FVS780-A"="Viability")

# keep original channel IDs, add antibody field
se <- annotate_antibodies(se, map)
```

## Save
```{r}
setwd(paste(working.dir, "1-data", sep = "/"))

saveRDS(se, "se.rds")
```

# Quality control

## Plot initial distributions
```{r}
dir.create(paste(working.dir, "2-quality-control", sep = "/"), showWarnings = FALSE)
setwd(paste(working.dir, "2-quality-control", sep = "/"))

pdf("quality-control.pdf",
    width = 10,
    height = 10)
(qc <- qc_flow_se_suite(se, assay_name = "exprs",
                       condition_col = "condition",
                       transform = "arcsinh", cofactor = 150,
                       label_field = "antibody"))
dev.off()
```

## Singlet Filter
  - gate singlets using Area~Height band method
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

res <- gate_singlets(
    se,
    method = "auto",        # AH = Area~Height band; AW = Area~Width one-sided
    assay_name = NULL,
    condition_col = "condition",
    label_field = "antibody",
    transform = "arcsinh",
    cofactor = 150,
    k = 3,                               # band width in MAD units
    per_condition = FALSE,
    max_points_per_condition = 50000,    # for plotting only
    seed = 1
)

pdf("singlets-filter.pdf",
    width = 5,
    height = 5)
res$plots
dev.off()

se.singlets <- res$se_filtered
```

## Debris Filter
  - Filter out debris based on FSC and SSC
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

res <- gate_debris_fsc_ssc(
  se.singlets,
  per_condition = FALSE,                 # global
  floor_mode = "quantile", floor_q = 0.06,  # 2% floors (tune 0.02–0.05)
  keep_fraction = 0.98,                  # ellipse for plotting/context
  keep_rule = "floors_or_ellipse",       # UNION: above floors OR inside ellipse
  transform = "arcsinh", cofactor = 150
)

pdf("debris-filter.pdf",
    width = 5,
    height = 5)
res$plots
dev.off()

se.singlet.debris <- res$se_filtered
```

## Viability filter
 - Keep cells within 3MAD of combined viability signal
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

# calculate MAD window 
res <- filter_cells_by_mad_window(se.singlet.debris, channel = "Viability", transform = "arcsinh", k = 4)
print(res$stats_overall)  # shows n_total, n_keep, pct_keep

# plot
pdf("viability-filter.pdf",
    width = 5,
    height = 5)
print(res$plot)
dev.off()

# apply filter
se.singlet.debris.viab <- res$se_filtered
```

## Outlier Filters

### H3
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

# calculate MAD window 
res <- filter_cells_by_mad_window(se.singlet.debris.viab, channel = "H3", transform = "arcsinh", k = 5)
print(res$stats_overall)  # shows n_total, n_keep, pct_keep

# plot
pdf("h3-filter.pdf",
    width = 5,
    height = 5)
print(res$plot)
dev.off()

# apply filter
se.singlet.debris.viab.outlier <- res$se_filtered
```

### H3K14
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

# calculate MAD window 
res <- filter_cells_by_mad_window(se.singlet.debris.viab.outlier, channel = "H3K14", transform = "arcsinh", k = 7)
print(res$stats_overall)  # shows n_total, n_keep, pct_keep

# plot
pdf("h3k14-filter.pdf",
    width = 5,
    height = 5)
print(res$plot)
dev.off()

# apply filter
se.singlet.debris.viab.outlier <- res$se_filtered
```

### H3K23
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

# calculate MAD window 
res <- filter_cells_by_mad_window(se.singlet.debris.viab.outlier, channel = "H3K23", transform = "arcsinh", k = 5)
print(res$stats_overall)  # shows n_total, n_keep, pct_keep

# plot
pdf("h3k23-filter.pdf",
    width = 5,
    height = 5)
print(res$plot)
dev.off()

# apply filter
se.singlet.debris.viab.outlier <- res$se_filtered
```

### Cell Cycle
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

# calculate MAD window 
res <- filter_cells_by_mad_window(se.singlet.debris.viab.outlier, channel = "Cell.Cycle", transform = "arcsinh", k = 5)
print(res$stats_overall)  # shows n_total, n_keep, pct_keep

# plot
pdf("cell.cycle-filter.pdf",
    width = 5,
    height = 5)
print(res$plot)
dev.off()

# apply filter
se.singlet.debris.viab.outlier <- res$se_filtered
```

## Plot filtered distributions
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

pdf("quality-control-filtered.pdf",
    width = 10,
    height = 10)
(qc <- qc_flow_se_suite(se.singlet.debris.viab.outlier, assay_name = "exprs",
                       condition_col = "condition",
                       transform = "arcsinh", cofactor = 150,
                       label_field = "antibody"))
dev.off()
```

## Save
```{r}
setwd(paste(working.dir, "2-quality-control", sep = "/"))

saveRDS(se.singlet.debris.viab.outlier, "se.singlet.debris.viab.outlier.rds")
```

# Stratify cells by cell cycle
  - Classify cells into cell cycle phases based on DNA content
  - Peaks from each condition are aligned before classification to account for global shifts in signal
```{r}
dir.create(paste(working.dir, "3-cell-cycle", sep = "/"), showWarnings = FALSE)
setwd(paste(working.dir, "3-cell-cycle", sep = "/"))

# classify cell cycle phases based on DNA content
res <- classify_cell_cycle_by_dna(
  se.singlet.debris.viab.outlier, channel = "Cell.Cycle",
  window_scope = "reference", reference_level = "Control",
  align_peaks = TRUE, align_scope = "reference",
  k = 1.5, min_s_frac = 0.1, transform = "none",
  g1_left_pad_mad = 3
)

table(res$se_with_phase$cell_cycle, res$se_with_phase$condition)

# plot cell cycle classification
pdf("cell-cycle-classification.pdf",
    width = 8,
    height = 5)
print(res$plot)
dev.off()

se.singlet.debris.viab.outlier.cc <- res$se_with_phase
```

## Save
```{r}
setwd(paste(working.dir, "3-cell-cycle", sep = "/"))

saveRDS(se.singlet.debris.viab.outlier.cc, "se.singlet.debris.viab.outlier.cc.rds")
```

# Differential abundance
```{r}
dir.create(paste(working.dir, "4-differential-abundance", sep = "/"), showWarnings = FALSE)
setwd(paste(working.dir, "4-differential-abundance", sep = "/"))

res <- compare_ptm_by_condition_phase(
  se.singlet.debris.viab.outlier.cc,
  features = c("H3K14ac","H3K23ac", "H3"),
  phase_col = "cell_cycle",
  assay_name = "exprs",
  transform_plot = "none",            # nicer for ratios
  sample_per_group = Inf,
  add_stat_labels = FALSE,
  drop_empty_facets = TRUE,
  facet_label_size = 15
)

pdf("unnomralized-differential-abundance.pdf",
    width = 7,
    height = 7)
res$plot
dev.off()

write.csv(res$stats, "unnormalized-differential-abundance-stats.csv", row.names = FALSE)
```

# PTM-PTM Covariation

### All cell cycle phases
```{r}
dir.create(paste(working.dir, "5-covariation", sep = "/"), showWarnings = FALSE)
setwd(paste(working.dir, "5-covariation", sep = "/"))

pdf("ptm-covariation-unnormalized-all-cellCycle.pdf",
    width = 7,
    height = 5)
res <- plot_ptm_covariation(
  se.singlet.debris.viab.outlier.cc.norm, 
  x_feature = "H3K14ac", 
  y_feature = "H3K23ac",
  assay_name = "exprs",
  transform = "none",
  sample_per_condition = Inf,
  outlier = "none", 
  outlier_q = c(0.01, 0.99),
  compute_stats_on = "filtered", 
  plot_filtered_only = TRUE,
  axis_label_prefix = '',
  title = "All cell cycle phases"
)
res$plot
dev.off()
```

### G0/G1 phase only
```{r}
setwd(paste(working.dir, "5-covariation", sep = "/"))

pdf("ptm-covariation-unnormalized-G0.G1-cellCycle.pdf",
    width = 7,
    height = 5)
res <- plot_ptm_covariation(
  se.singlet.debris.viab.outlier.cc.norm[,se.singlet.debris.viab.outlier.cc.norm$cell_cycle == "G0/G1"],
  x_feature = "H3K14ac", 
  y_feature = "H3K23ac",
  assay_name = "exprs",
  transform = "none",
  sample_per_condition = Inf,
  outlier = "none", 
  outlier_q = c(0.01, 0.99),
  compute_stats_on = "filtered", 
  plot_filtered_only = TRUE,
  axis_label_prefix = '',
  title = "G0/G1 phase"
)
res$plot
dev.off()
```

### S phase only
```{r}
setwd(paste(working.dir, "5-covariation", sep = "/"))

pdf("ptm-covariation-unnormalized-S-cellCycle.pdf",
    width = 7,
    height = 5)
res <- plot_ptm_covariation(
  se.singlet.debris.viab.outlier.cc.norm[,se.singlet.debris.viab.outlier.cc.norm$cell_cycle == "S"],
  x_feature = "H3K14ac", 
  y_feature = "H3K23ac",
  assay_name = "exprs",
  transform = "none",
  sample_per_condition = Inf,
  outlier = "none", 
  outlier_q = c(0.01, 0.99),
  compute_stats_on = "filtered", 
  plot_filtered_only = TRUE,
  axis_label_prefix = '',
  title = "S phase"
)
res$plot
dev.off()
```

### G2/M phase only
```{r}
setwd(paste(working.dir, "5-covariation", sep = "/"))

pdf("ptm-covariation-unnormalized-G2.M-cellCycle.pdf",
    width = 7,
    height = 5)
res <- plot_ptm_covariation(
  se.singlet.debris.viab.outlier.cc.norm[,se.singlet.debris.viab.outlier.cc.norm$cell_cycle == "G2/M"],
  x_feature = "H3K14ac", 
  y_feature = "H3K23ac",
  assay_name = "exprs",
  transform = "none",
  sample_per_condition = Inf,
  outlier = "none", 
  outlier_q = c(0.01, 0.99),
  compute_stats_on = "filtered", 
  plot_filtered_only = TRUE,
  axis_label_prefix = '',
  title = "G2/M phase"
)
res$plot
dev.off()
```